\pagestyle{empty}
\pagenumbering{roman}

% -----
% TITLE PAGE
% -----
\begin{titlepage}
  \begin{center}
    %\vspace*{1.0cm}

    \Huge
    {\bf Deterministic and Probabilistic \\ Bijective Combinatorics \\ for Macdonald Polynomials} \\

    \vspace*{1.0cm}

    \normalsize
    by \\

    \vspace*{1.0cm}

    \Large
    Guilherme Zeus Dantas e Moura \\

    \vspace*{3.0cm}

    \normalsize
    A thesis \\
    presented to the University of Waterloo \\
    in fulfillment of the \\
    thesis requirement for the degree of \\
    Master of Mathematics \\
    in \\
    Combinatorics and Optimization \\

    \vspace*{2.0cm}

    Waterloo, Ontario, Canada, 2025 \\

    \vspace*{1.0cm}

    \copyright\ Guilherme Zeus Dantas e Moura 2025 \\
  \end{center}
\end{titlepage}

\pagestyle{plain}
\setcounter{page}{2}

% -----
% LIST OF TODO NOTES
% -----
% \listoftodos
% 
\cleardoublepage
\phantomsection

% -----
% AUTHOR'S DECLARATION
% -----

{\large
%\addcontentsline{toc}{chapter}{Author's Declaration}
\begin{center}\textbf{Author's Declaration}\end{center}

This thesis consists of material all of which I authored or co-authored: see \hyperref[sec:contributions]{Statement of Contributions} included in the thesis.
% This is a true copy of the thesis, including any required final revisions, as accepted by my examiners.
This is the version of the thesis sent to the examiners, not including any corrections or revisions from their comments.
\bigskip

I understand that my thesis may be made electronically available to the public.

}

\cleardoublepage
\phantomsection

% -----
% STATEMENT OF CONTRIBUTIONS
% -----
\label{sec:contributions}

{\large
%\addcontentsline{toc}{chapter}{Statement of Contributions}
\begin{center}\textbf{Statement of Contributions}\end{center}

I am the sole author of \Cref{chap:introduction,chap:polynomials,chap:models}, which were written under the supervision of Olya Mandelshtam.

The research presented in \Cref{chap:bijective} is based on joint work with Olya Mandelshtam, to appear in an upcoming publication.

The research presented in \Cref{chap:probabilistic} is based on joint work with Olya Mandelshtam, in \cite{DM25}:
\fullcite{DM25}.
% This is a true copy of the thesis, including any required final revisions, as accepted by my examiners.

}

\cleardoublepage
\phantomsection

% -----
% ABSTRACT
% -----

{\large
%\addcontentsline{toc}{chapter}{Abstract}
\begin{center}\textbf{Abstract}\end{center}

Permuted-basement Macdonald polynomials $E_\alpha^\sigma(\mathbf{x};q,t)$ are nonsymmetric generalizations of symmetric Macdonald polynomials indexed by a composition $\alpha$ and a permutation $\sigma$.
They form a basis for the polynomial ring $\rationals(q,t)[\mathbf{x}]$ for each fixed permutation $\sigma$.
They can be described combinatorially as generating functions over augmented fillings of composition shape $\alpha$ with a basement permutation $\sigma$.

We construct (deterministic) bijections and probabilistic bijections on fillings that
prove identities for the relationship between $E_\alpha^\sigma$, $E_\alpha^{\sigma s_i}$, $E_{s_i \permact \alpha}^\sigma$, and $E_{s_i \permact \alpha}^{\sigma s_i}$.
These identities correspond to two combinatorial operations on the shape and basement of the fillings:
swapping adjacent entries in the basement, generalizing a result of \textcite{Ale19};
and swapping adjacent parts in the shape, giving a formula for expanding $E_\alpha^\sigma$ in terms of $E_{s_i \permact \alpha}^\sigma$ and $E_{s_i \permact \alpha}^{\sigma s_i}$.

}

\cleardoublepage
\phantomsection

% -----
% ACKNOWLEDGEMENTS
% -----

{\large
%\addcontentsline{toc}{chapter}{Acknowledgements}
\begin{center}\textbf{Acknowledgements}\end{center}

% This research was partially supported by NSERC grant RGPIN-2021-02568.

I would like to thank my supervisor, Olya Mandelshtam, for her close collaboration throughout my Master's program.
It is often said that choosing a supervisor is one of the most impactful decisions a graduate student makes,
and I am extremely happy with and grateful for the choice I have made.

I thank the members of my thesis committee, Oliver Pechenik and Stephan Pfannerer-Mittas, for their time and valuable feedback.
I extend these thanks to all the professors who taught me at the University of Waterloo,
and to the faculty of the Algebraic Combinatorics group.
The supportive environment they have cultivated has been essential to my success as a graduate student.
I also thank the staff of the Combinatorics and Optimization department for their prompt and helpful assistance.

I am grateful for the friends I made in Waterloo and look forward to growing these friendships in the years to come.
I thank my parents for all their support and for the freedom that they granted me to explore any one of my interests, even when it has taken me far from home.

Finally, the writing and research of this thesis was only possible because of
the 24-hour access to my office in the Mathematics \& Computer Building, 
the ability to safely walk home at extremely late hours,
the cheap coffee at the Mathematics Coffee \& Donuts Shop,
the kind workers at Williams Fresh Cafe and at Roosters Fried Chicken,
and many cans of sugarfree Red Bull.

}

\cleardoublepage
\phantomsection

% -----
% TABLE OF CONTENTS
% -----
\renewcommand\contentsname{Table of Contents}
\tableofcontents
\cleardoublepage
\phantomsection

% -----
% LIST OF FIGURES
% -----
%\addcontentsline{toc}{chapter}{List of Figures}
\listoffigures
\cleardoublepage
\phantomsection

% -----
% LIST OF TABLES
% -----
% \addcontentsline{toc}{chapter}{List of Tables}
% \listoftables
% \cleardoublepage
% \phantomsection

% -----
% LIST OF ABBREVIATIONS
% -----
% \renewcommand*{\abbreviationsname}{List of Abbreviations}
% \printglossary[type=abbreviations]
% \cleardoublepage
% \phantomsection

% -----
% LIST OF SYMBOLS
% -----
% \printglossary[type=symbols]
% \cleardoublepage
% \phantomsection

\pagenumbering{arabic}