\newcommand{\coinphi}{%
  \tikz[baseline=(char.base)]{
    \node[scale=.95,shape=circle,fill=yellow!85!orange!70,draw=yellow!85!orange!70!black,inner sep=1pt] (char) {\textcolor{black}{$\phi$}};
  }%
}

\newcommand{\coinpsi}{%
  \tikz[baseline=(char.base)]{
    \node[scale=.95,shape=circle,fill=gray!50!blue!30,draw=gray!50!blue!30!black!50,inner sep=1pt] (char) {\textcolor{black}{$\psi$}};
  }%
}

\newcommand{\labelS}{\tikz[baseline=(char.base)]{
    \node[draw=green!70!black, fill=green!20, rounded corners=2pt, inner sep=2pt] (char) {\texttt{S}};
}}
\newcommand{\labelN}{\tikz[baseline=(char.base)]{
    \node[draw=red!70!black, fill=red!20, rounded corners=2pt, inner sep=2pt] (char) {\texttt{N}};
}}
\newcommand{\labelC}{\tikz[baseline=(char.base)]{
    \node[draw=orange!80!black, fill=orange!20, rounded corners=2pt, inner sep=2pt] (char) {\texttt{C}};
}}

\chapter{Bijective combinatorics for Macdonald polynomials} \label{chap:bijective}

\todo{Write all of the polynomial identities in the introduction of this chapter.}

In the first part of this chapter,
we study a rearrangement property of the modified Macdonald polynomials \(\tilde{H}_\mu(\mathbf{x}; q, t)\),
\Cref{thm:tildeH-rearrangement},
that for any partition \(\mu\) and any rearrangement \(\alpha\) of \(\mu\),
\begin{equation} \label{eq:tildeH-rearrangement}
    \sum_{T \in \Fill[\positives](\mu)} q^{\inv(T)} t^{\maj(T)} \mathbf{x}^T
    =
    \sum_{T \in \Fill[\positives](\alpha)} q^{\inv(T)} t^{\maj(T)} \mathbf{x}^T.
\end{equation}
The sum over \(\Fill[\positives](\mu)\) is the celebrated combinatorial formula for \(\tilde{H}_\mu(\mathbf{x}; q, t)\) from \cite{HHL05} (see \Cref{subsec:modified-macdonald}).
In \cite{HHL08}, the rearrangement property was established by proving that the sum over \(T \in \Fill[\positives](\mu)\) is equal to \(\tilde{H}_\mu(\mathbf{x}; q, t)\), essentially copying the argument from \cite{HHL05} to the new setting.

Here, we present a direct bijective proof of \Cref{eq:tildeH-rearrangement}, without going through \(\tilde{H}_\mu(\mathbf{x}; q, t)\). 
The key step is \Cref{thm:phi-maj-inv},
that guarantees the existence of a bijection \(\phi \colon \Fill(\alpha) \to \Fill(s_i \permact \alpha)\) that preserves \(\content\), \(\inv\), and \(\maj\).

We present two constructions of the bijection \(\phi\).
The first construction is something with blocks,
and the second construction is an iterative algorithm from \cite{KLO22}, which we present in a clearer and more systematic form than the original exposition.
Each construction has its own advantages and disadvantages, and we present both for completeness.
We prove that both algorithms coincide in ...,
and then prove \Cref{thm:phi-maj-inv} in two different ways,
one for each construction.
\todo{There are ``two levels'' in which I am making improvements from \cite{KLO22}: the first is finding an entirely new way of getting to the bijection (through blocks), and the second is rewriting their algorithm in a clearer way. I feel like these two different ``improvements'' are getting confused in this sentence.}

An auxiliary bijection \(\psi \colon \Fill[\mathcal{A}](\alpha) \to \Fill[\mathcal{A}](s_i \permact \alpha)\) is also defined, that preserves \(\content\) but changes \(\inv\) and \(\maj\) in a controlled way, as specified in \Cref{thm:psi-maj-inv}.

In the second part of this chapter,
we prove \Cref{thm:three-term-a-non-integral,thm:three-term-b-non-integral},
pair of three-term identities for permuted-basement Macdonald polynomials \(E_\alpha^\sigma(\mathbf{x}; q, t)\), \(E_{s_i \permact \alpha}^\sigma(\mathbf{x}; q, t)\), and \(E_{s_i \permact \alpha}^{\sigma s_i}(\mathbf{x}; q, t)\).

We use the same bijection \(\phi\) on non-augmented fillings (as well as an auxiliary bijection \(\psi\)) as a building block for a bijection on augmented fillings.

\section{Bijections on non-augmented fillings and the rearrangement property of \texorpdfstring{\(\tilde{H}_\mu\)}{H̃μ}} \label{sec:bijective-non-augmented}

In this section, we define the bijection \(\phi \colon \Fill(\alpha) \to \Fill(s_i \permact \alpha)\) from \cite{KLO22} that preserves the statistics \(\inv\) and \(\maj\).
As an auxiliary construction,
we also define another bijection \(\psi \colon \Fill(\alpha) \to \Fill(s_i \permact \alpha)\) that changes the statistics \(\inv\) and \(\maj\) in a controlled way.
The manner in which \(\psi\) changes the statistics is specified in \Cref{thm:psi-maj-inv}.
This auxiliary bijection \(\psi\) will also be important in a later section.
Without loss of generality, we restrict to the case \(\alpha_{i} > \alpha_{i+1}\).

Both of \(\phi\) and \(\psi\) follow a very rigid structure.
If a non-augmented filling \(T \in \Fill(\alpha)\) is mapped to a filling \(U \in \Fill(s_i \permact \alpha)\) by one of these maps, then:
\begin{itemize}
    \item all columns except \(i\) and \(i+1\) are identical in \(T\) and \(U\), that is,
    \begin{equation} \label{eq:restriction-other-columns}
        T(j, r) = U(j, r) \quad \text{for all } j \in \interval{n} \setminus \{i, i+1\} \text{ and all } r \in \interval{\alpha_j};
    \end{equation}
    \item the topmost \(\alpha_i - \alpha_{i+1}\) entries in column \(i\) of \(T\) are copied to the topmost \(\alpha_i - \alpha_{i+1}\) entries in column \(i+1\) of \(U\), that is,
    \begin{equation} \label{eq:restriction-top-entries}
        T(i, r) = U(i+1, r) \quad \text{for all } r \in \interval{\alpha_{i+1} + 1, \alpha_i};
    \end{equation}
    \item the entries in rows \(1, \dots, \alpha_{i+1}\) of columns \(i\) and \(i+1\) in \(T\) and \(U\) are the same, but they might be in the same positions or in swapped positions, that is,
    \begin{equation} \label{eq:restriction-swap-per-row}
        \{ T(i, r), T(i+1, r) \} = \{ U(i, r), U(i+1, r) \} \quad \text{for all } r \in \interval{\alpha_{i+1}}.
    \end{equation}
\end{itemize}

Given these restrictions, the only remaining freedom in determining \(\phi\) and \(\psi\) is to decide, for each \(r \in \interval{\alpha_{i+1}}\), whether to set
\begin{equation}
    \big(U(i,r), U(i+1,r)\big) = \big(T(i,r), T(i+1,r)\big) \text{ or } \big(U(i,r), U(i+1,r)\big) = \big(T(i+1,r), T(i,r)\big).
\end{equation}
In the first case, we say the entries of \(T\) and \(U\) in row \(r\) of columns \(i\) and \(i+1\) are \vocab{not swapped};
in the second case, we say they are \vocab{swapped}.
Below, we present two different algorithms to decide, for each \(r \in \interval{\alpha_{i+1}}\), whether the entries in row \(r\) of columns \(i\) and \(i+1\) are swapped or not swapped.

\subsection{Block structure algorithm} \label{subsec:block-structure}

\begin{proposition} \label{prop:odd-ind-magic}
    Let \(a, b, c, d \in \mathcal{A}\) be four elements of an alphabet \(\mathcal{A}\).
    The following statements are equivalent:
    \begin{enumerate}
        \item \label{item:odd-ind}
        \(\ind{a, d} + \ind{b, c} - \ind{a, c} - \ind{b, d} = \pm 1\);
        \item \label{item:triple-equal-ind}
        \(\ind{a, b, c} = \ind{b, c, d} = \ind{a, d, c}\) and \(a \neq b\) and \(c \neq d\);
    \end{enumerate}
\end{proposition}

\begin{proof}
    It suffices to check all possible relative orders of \(a, b, c, d\) (including equalities).
    Computational verification shows that the two statements are equivalent.
    \todo{Attempt find a neater proof.}
\end{proof}

Let \(r \in \interval{\alpha_{i+1} - 1}\).
Let \( a = T(i,r+1)\), \(b = T(i+1,r+1)\), \(c = T(i,r)\), and \(d = T(i+1,r)\).
We say that rows \(r\) and \(r + 1\) of \(T\) are \vocab{intertwined} if \(a, b, c, d\) satisfy the equivalent conditions of \Cref{prop:odd-ind-magic}.
Otherwise, we say that rows \(r\) and \(r + 1\) of \(T\) are \vocab{not intertwined}.

The interval of rows \(\interval{\alpha_{i+1}}\) is uniquely partitioned into intervals, called \vocab{blocks}, such that two consecutive rows in the same block are intertwined, and two consecutive rows in different blocks are not intertwined.
\Cref{fig:blocks} shows a filling and its blocks.
 
\begin{figure}[htbp]
    \centering
	\begin{tikzpicture}[scale=.5]
		\node at (-0.5, 7.5) {\tiny \(7\)};
		\node at (-0.5, 6.5) {\tiny \(6\)};
		\node at (-0.5, 5.5) {\tiny \(5\)};
		\node at (-0.5, 4.5) {\tiny \(4\)};
		\node at (-0.5, 3.5) {\tiny \(3\)};
		\node at (-0.5, 2.5) {\tiny \(2\)};
		\node at (-0.5, 1.5) {\tiny \(1\)};
		\node[anchor=north] at (0.5, 1) {\(\scriptstyle i\)};
		\node[anchor=north] at (1.5, 1) {\(\scriptstyle i+1\)};
		\draw (0, 7) rectangle ++(1, 1);
		\draw (1, 6) rectangle ++(1, 1)[fill=gray!50];
		\draw (0, 6) rectangle ++(1, 1)[fill=gray!50];
		\draw (1, 5) rectangle ++(1, 1)[fill=gray!50];
		\draw (0, 5) rectangle ++(1, 1)[fill=gray!50];
		\draw (1, 4) rectangle ++(1, 1);
		\draw (0, 4) rectangle ++(1, 1);
		\draw (1, 3) rectangle ++(1, 1);
		\draw (0, 3) rectangle ++(1, 1);
		\draw (1, 2) rectangle ++(1, 1);
		\draw (0, 2) rectangle ++(1, 1);
		\draw (1, 1) rectangle ++(1, 1)[fill=gray!50];
		\draw (0, 1) rectangle ++(1, 1)[fill=gray!50];
		\node at (0.5, 7.5) {\(3\)};
		\node at (1.5, 6.5) {\(5\)};
		\node at (0.5, 6.5) {\(1\)};
		\node at (1.5, 5.5) {\(2\)};
		\node at (0.5, 5.5) {\(4\)};
		\node at (1.5, 4.5) {\(8\)};
		\node at (0.5, 4.5) {\(5\)};
		\node at (1.5, 3.5) {\(6\)};
		\node at (0.5, 3.5) {\(1\)};
		\node at (1.5, 2.5) {\(7\)};
		\node at (0.5, 2.5) {\(3\)};
		\node at (1.5, 1.5) {\(6\)};
		\node at (0.5, 1.5) {\(4\)};
		\draw[decorate, decoration={brace, amplitude=5pt, mirror, raise=5pt}] (2, 5.1) -- (2, 6.9) node[midway, anchor=west, xshift=10pt] {third block};
		\draw[decorate, decoration={brace, amplitude=5pt, mirror, raise=5pt}] (2, 2.1) -- (2, 4.9) node[midway, anchor=west, xshift=10pt] {second block};
		\draw[decorate, decoration={brace, amplitude=5pt, mirror, raise=5pt}] (2, 1.1) -- (2, 1.9) node[midway, anchor=west, xshift=10pt] {first block};
	\end{tikzpicture}
	\caption{Columns \(i\) and \(i+1\) of a filling \(T\) with three blocks.}
	\label{fig:blocks}
\end{figure}

Let's interpret the intertwining condition via~\Cref{item:triple-equal-ind} of \Cref{prop:odd-ind-magic}.
The number \(\ind{a, b, c}\) is the indicator of whether the triple \((i, r+1), (i+1, r+1), (i, r)\) is a Type A inversion triple in \(T\).
The number \(\ind{b, c, d}\) is the indicator of whether the triple \((i+1, r+1), (i, r), (i+1, r)\) is a Type B inversion triple in \(U\) if rows \(r\) and \(r + 1\) are not swapped.
The number \(\ind{a, d, c}\) is the indicator of whether the triple \((i, r+1), (i+1, r), (i, r)\) is a Type B inversion triple in \(U\) if rows \(r\) and \(r + 1\) are swapped.

Therefore,
if we require that rows in the same block are either all swapped or all not swapped in \(U\),
then we guarantee that the number of inversion triples within each block is the same in \(T\) and \(U\), regardless of whether the block is swapped or not swapped.

The remaining freedom is to decide, for each block, whether to swap it or not swap it.
We use this freedom to attempt to guarantee that the number of inversion triples between different blocks is also the same in \(T\) and \(U\).

Consider a block of \(T\).
Let \(a\) and \(b\) be the entries in columns \(i\) and \(i+1\) of the top row of the block, respectively.
Let \(c\) and \(d\) be the entries in columns \(i\) and \(i+1\) of the bottom row of the block, respectively.
Let \(e_1\) and \(e_2\) be the entries in columns \(i\) and \(i+1\) of the row immediately above the block, respectively.
In the case that the block is the topmost block, we leave \(e_2\) undefined.
Let \(f_1\) and \(f_2\) be the entries in columns \(i\) and \(i+1\) of the row immediately below the block, respectively.
In the case that the block is the bottommost block, we take \(f_1 = f_2 = \infty\).
 
\begin{figure}[htbp]
    \centering
	\begin{tikzpicture}[scale=.5]
		\draw (0, 0) rectangle ++(1, 1);
		\draw (1, 0) rectangle ++(1, 1);
		\draw[fill=gray!50] (0, 1) rectangle ++(1, 1);
		\draw[fill=gray!50] (1, 1) rectangle ++(1, 1);
		\fill[gray!50] (0, 2) rectangle ++(2, 2);
		\node at (1, 3.25) {\(\vdots\)};
		\draw[fill=gray!50] (0, 4) rectangle ++(1, 1);
		\draw[fill=gray!50] (1, 4) rectangle ++(1, 1);
		\draw (0, 5) rectangle ++(1, 1);
		\draw (1, 5) rectangle ++(1, 1);
		\node at (0.5, 5.5) {\(e_1\)};
		\node at (1.5, 5.5) {\(e_2\)};
		\node at (0.5, 4.5) {\(a\)};
		\node at (1.5, 4.5) {\(b\)};
		\node at (0.5, 1.5) {\(c\)};
		\node at (1.5, 1.5) {\(d\)};
		\node at (0.5, 0.5) {\(f_1\)};
		\node at (1.5, 0.5) {\(f_2\)};
	\end{tikzpicture}
	\caption{A block of a filling \(T\), with the entries \(a, b, c, d, e_1, e_2, f_1, f_2\) indicated.}
	\label{fig:general-block}
\end{figure}

Since the rows containing \(f_1, f_2\) and \(c, d\) are not intertwined, \(\ind{c, d, f_1} = \ind{c, d, f_2}\).
We denote this common value by \(\ind{c, d, f_\bullet}\).
Similarly, since the rows containing \(a, b\) and \(e_1, e_2\) are not intertwined, \(\ind{e_1, a, b} = \ind{e_2, a, b}\).
We denote this common value by \(\ind{e_\bullet, a, b}\).

\begin{definition}[The bijection \(\phi\) via blocks]
    Let \(T \in \Fill(\alpha)\).
    We define \(U = \phi(T) \in \Fill(s_i \permact \alpha)\)
    by requiring \Cref{eq:restriction-other-columns,eq:restriction-top-entries,eq:restriction-swap-per-row} and requiring that, for each block of \(T\),
    the rows of the block are swapped in \(U\) if and only if \(\ind{c, d, f_\bullet} \neq \ind{e_\bullet, a, b}\),
    where \(a, b, c, d, e_\bullet, f_\bullet\) are as defined above.
\end{definition}

\begin{definition}[The bijection \(\psi\) via blocks]
    Let \(T \in \Fill(\alpha)\).
    We define \(U = \psi(T) \in \Fill(s_i \permact \alpha)\)
    by requiring \Cref{eq:restriction-other-columns,eq:restriction-top-entries,eq:restriction-swap-per-row} and requiring that, for each block of \(T\) except the bottommost one,
    the rows of the block are swapped in \(U\) if and only if \(\ind{c, d, f_\bullet} \neq \ind{e_\bullet, a, b}\),
    and that the rows of the bottommost block are swapped in \(U\) if and only if \(\ind{c, d, f_\bullet} = \ind{e_\bullet, a, b}\),
    where \(a, b, c, d, e_\bullet, f_\bullet\) are as defined above.
\end{definition}

\Cref{fig:output-filling} shows the image of the filling in \Cref{fig:blocks} under the map \(\phi\), following this definition.

\begin{figure}[htbp]
    \centering
	\begin{tikzpicture}[scale=.5]
		\node at (-0.5, 7.5) {\tiny \(7\)};
		\node at (-0.5, 6.5) {\tiny \(6\)};
		\node at (-0.5, 5.5) {\tiny \(5\)};
		\node at (-0.5, 4.5) {\tiny \(4\)};
		\node at (-0.5, 3.5) {\tiny \(3\)};
		\node at (-0.5, 2.5) {\tiny \(2\)};
		\node at (-0.5, 1.5) {\tiny \(1\)};
		\node[anchor=north] at (0.5, 1) {\(\scriptstyle i\)};
		\node[anchor=north] at (1.5, 1) {\(\scriptstyle i+1\)};
		\draw (1, 7) rectangle ++(1, 1);
		\draw (1, 6) rectangle ++(1, 1)[fill=gray!50];
		\draw (0, 6) rectangle ++(1, 1)[fill=gray!50];
		\draw (1, 5) rectangle ++(1, 1)[fill=gray!50];
		\draw (0, 5) rectangle ++(1, 1)[fill=gray!50];
		\draw (1, 4) rectangle ++(1, 1);
		\draw (0, 4) rectangle ++(1, 1);
		\draw (1, 3) rectangle ++(1, 1);
		\draw (0, 3) rectangle ++(1, 1);
		\draw (1, 2) rectangle ++(1, 1);
		\draw (0, 2) rectangle ++(1, 1);
		\draw (1, 1) rectangle ++(1, 1)[fill=gray!50];
		\draw (0, 1) rectangle ++(1, 1)[fill=gray!50];
		\node at (1.5, 7.5) {\(3\)};
		\node at (1.5, 6.5) {\(5\)};
		\node at (0.5, 6.5) {\(1\)};
		\node at (1.5, 5.5) {\(2\)};
		\node at (0.5, 5.5) {\(4\)};
		\node at (1.5, 4.5) {\(5\)};
		\node at (0.5, 4.5) {\(8\)};
		\node at (1.5, 3.5) {\(1\)};
		\node at (0.5, 3.5) {\(6\)};
		\node at (1.5, 2.5) {\(3\)};
		\node at (0.5, 2.5) {\(7\)};
		\node at (1.5, 1.5) {\(6\)};
		\node at (0.5, 1.5) {\(4\)};
		\draw[decorate, decoration={brace, amplitude=5pt, mirror, raise=5pt}] (2, 5.1) -- (2, 6.9) node[midway, anchor=west, xshift=10pt] {unchanged, since \(\ind{4, 2, 5} = 1 = \ind{3, 1, 5}\)};
		\draw[decorate, decoration={brace, amplitude=5pt, mirror, raise=5pt}] (2, 2.1) -- (2, 4.9) node[midway, anchor=west, xshift=10pt] {swapped, since \(\ind{2, 5, 8} = 0 \neq 1 = \ind{3, 7, 4}\)};
		\draw[decorate, decoration={brace, amplitude=5pt, mirror, raise=5pt}] (2, 1.1) -- (2, 1.9) node[midway, anchor=west, xshift=10pt] {unchanged, since \(\ind{4, 6, \infty} = 0 = \ind{7, 4, 6}\)};
	\end{tikzpicture}
	\caption{Columns \(i\) and \(i+1\) of the filling \(U\) derived from the filling \(T\) in \cref{fig:blocks}.}
	\label{fig:output-filling}
\end{figure}

In \Cref{prop:phi-psi-bijections,thm:phi-maj-inv}, we prove that \(\phi\) is a bijection that preserves \(\content\), \(\inv\), and \(\maj\).
Let's highlight the parts that are straightforward from the definition via blocks.

It is straightforward to verify that both \(\phi\) and \(\psi\) are bijections, since the the block structure of \(T\) is the same as that of \(U = \phi(T)\) or \(U = \psi(T)\), and the swapping condition can be reversed to recover \(T\) from \(U\).
\Cref{eq:restriction-other-columns,eq:restriction-top-entries,eq:restriction-swap-per-row} guarantee that \(\content(T) = \content(U)\).

Moreover, it is straightforward to verify that the number of inversion triples within columns \(i\) and \(i+1\) is the same in \(T\) and \(\phi(T)\), since the number of inversion triples within each block is preserved, and the choice of swapping or not swapping each block is made precisely to ensure that the number of inversion triples between different blocks is also preserved.
It is not hard to see that this implies that \(\inv(T) = \inv(\phi(T))\).

The fact that \(\maj(T) = \maj(\phi(T))\) is less straightforward,
since we haven't made any direct attempt to control the major index in our definition of \(\phi\).
We prove this in \Cref{thm:phi-maj-inv}.

\subsection{Iterative algorithm}
Let \(T \in \Fill(\alpha)\) with \(\alpha_i > \alpha_{i+1}\).
We describe the algorithm to compute \(U = \phi(T)\) or \(U = \psi(T)\).
The algorithm uses a \vocab{coin} with two sides, \(\coinphi\) and \(\coinpsi\), as a bookkeeping device.
Each row from \(1\) to \(\alpha_{i+1} - 1\) is, at some point of the algorithm, assigned a \vocab{coin value} in \(\{\coinphi, \coinpsi\}\).
Initially, assign the coin value of row \(1\) to be \(\coinphi\) if computing \(U = \phi(T)\), or \(\coinpsi\) if computing \(U = \psi(T)\).

In addition to the coin value, each row from \(1\) to \(\alpha_{i+1}\) is labeled from \(\{\labelN, \labelS, \labelC\}\).
A label of \(\labelN\) means the entries in row \(r\) of columns \(i\) and \(i+1\) in \(T\) and \(U\) are \emph{not} swapped;
a label of \(\labelS\) means they are \emph{swapped};
and a label of \(\labelC\) indicates that the behavior for row \(r\) should \emph{copy} the behavior for the row above it.

We process the rows iteratively from row \(1\) up to row \(\alpha_i\).
For each \(r \in \{1,2,\dots,\alpha_{i+1}-1\}\),
let \( a = T(i,r+1)\), \(b = T(i+1,r+1)\), \(c = T(i,r)\), and \(d = T(i+1,r)\).
Note that \(\ind{a, d} = \ind{b, c} \neq \ind{a, c} = \ind{b, d}\) is impossible by transitivity,
thus \( \ind{a,d} + \ind{b,c} - \ind{a,c} - \ind{b,d} \in \{1, 0, -1\} \).

\begin{description}
    \item[Case I.] If 
    \(\ind{a,d} + \ind{b,c} - \ind{a,c} - \ind{b,d} = \pm 1\),
    (cases 3,4,9,10 in \cite{KLO22}),  
    then assign row \(r\) the label \(\labelC\).
    Assign row \(r+1\) the coin value of row \(r\) if \(\chi_{ab} = \chi_{cd}\) (cases 3,9),
    and assign row \(r+1\) the opposite coin value of row \(r\) if \(\chi_{ab} \neq \chi_{cd}\) (cases 4,10).

    \item[Case II.] If
    \(\ind{a,c} = \ind{a,d} = \ind{b,c} = \ind{b,d}\)
    (cases 1,2,11,12),
    then label row \(r\) with \(\labelN\) (resp.~\(\labelS\)) if its coin value is \(\coinphi\) (resp.~\(\coinpsi\)).
    Assign row \(r+1\) the coin value \(\coinphi\).
    
    \item[Case III.] If 
    \(\ind{a,c} = \ind{b,c} \neq \ind{a,d} = \ind{b,d}\)  
    (cases 7,8),  
    then label row \(r\) with \(\labelS\) (resp.~\(\labelN\)) if its coin value is \(\coinphi\) (resp.~\(\coinpsi\)).
    Assign row \(r+1\) the coin value \(\coinphi\).
    
    \item[Case IV.] If
    \(\ind{a,c} = \ind{a,d} \neq \ind{b,c} = \ind{b,d}\)
    (cases 5,6),
    then label row \(r\) with \(\labelN\) (resp.~\(\labelS\)) if its coin value is \(\coinphi\) (resp.~\(\coinpsi\)).
    Assign row \(r+1\) the coin value \(\coinpsi\).
\end{description}

Finally, we process row \(r = \alpha_{i+1}\) (the last row of column \(i+1\)) as follows.
Let \( a = T(i,r+1)\), \(c = T(i,r)\), and \(d = T(i+1,r)\).

\begin{description}
    \item[Base I.] If \(\ind{a,c} \neq \ind{a,d}\), then label row \(r\) with
    \(\labelS\) (resp.~\(\labelN\)) if its coin value is \(\coinphi\) (resp.~\(\coinpsi\)).

    \item[Base II.] If \(\ind{a,c} = \ind{a,d}\), then label row \(r\) with
    \(\labelN\) (resp.~\(\labelS\)) if its coin value is \(\coinphi\) (resp.~\(\coinpsi\)).
\end{description}

\todo{There's a way to compactify these cases into overlapping conditions, which would also encompass the base cases.}

Finally, build \(U\) row by row:
\begin{itemize}
    \item If row \(r\) is labeled \(\labelN\), set
    \(U(i,r) = T(i,r)\) and \(U(i+1,r) = T(i+1,r)\).
    \item If row \(r\) is labeled \(\labelS\), set
    \(U(i,r) = T(i+1,r)\) and \(U(i+1,r) = T(i,r)\).
    \item If row \(r\) is labeled \(\labelC\), then use the same rule as the nearest row above it labeled \(\labelN\) or \(\labelS\).
    Note that since row \(\alpha_{i+1}\) is always labeled \(\labelN\) or \(\labelS\), there is always a nearest row above it labeled \(\labelN\) or \(\labelS\).
\end{itemize}

\todo{Add an example of the iterative algorithm.}

\subsection{Recursive algorithm}
The algorithm above may be described recursively in terms of the labels, instead of iteratively.

Let \(T \in \Fill(\alpha)\) with \(\alpha_i > \alpha_{i+1}\).
Assign the coin value of row \(1\) to be \(\coinphi\) if computing \(U = \phi(T)\), or \(\coinpsi\) if computing \(U = \psi(T)\).
If column \(i+1\) consists of a single entry, then assign the label of this row according to Base~I or~II.
Otherwise, assign the label of row \(1\) according one of Cases~I--IV,
and at the same time determine the coin value to be used for the next row.
Remove the first row to obtain the sub-filling \(T'\), and apply the algorithm recursively to \(T'\) with this new coin value, obtaining the labels for rows \(2,\dots,\alpha_{i+1}\).
Once all labels are determined, construct \(U\) in a final pass by applying the rules for \(\labelN,\labelS,\labelC\) as before.

It is straightforward to verify that both the iterative and recursive descriptions of the algorithm are equivalent.
This recursive interpretation is particularly useful for proving properties about the bijections \(\phi\) and \(\psi\), as it allows us to use induction on the number of rows in the filling.

\todo{Add an example of the recursive algorithm.}


\subsection{Properties of the bijections \texorpdfstring{\(\phi\)}{𝜙} and \texorpdfstring{\(\psi\)}{𝜓}}

\begin{proposition} \label{prop:phi-psi-bijections}
    The maps \(\phi\) and \(\psi\) are bijections.
\end{proposition}

\begin{theorem}[{\cite[Proposition~4.3]{KLO22}}] \label{thm:phi-maj-inv}
    The bijection \(\phi\) preserves the statistics \(\inv\) and \(\maj\); that is, for all \(T \in \Fill(\alpha)\), if \(U = \phi(T)\), then
    \begin{equation}
        \inv(T) = \inv(U) \quad \text{and} \quad \maj(T) = \maj(U).
    \end{equation}
\end{theorem}

\begin{theorem}[{\cite[14]{KLO22}}] \label{thm:psi-maj-inv}
    The bijection \(\psi\) modifies the statistics \(\inv\) and \(\maj\) in a specific way. For any \(T \in \Fill(\alpha)\) and \(U = \psi(T)\), we have
    \begin{itemize}
        \item if \(\ind{{T(i,1)},{T(i+1,1)}} = 1\) and \(\ind{{U(i,1)},{U(i+1,1)}} = 1\), then
        \begin{equation}
            \maj(T) = \maj(U) - (L+1)
            \quad \text{and} \quad
            \inv(T) = \inv(U) + A,
        \end{equation}
        \item if \(\ind{{T(i,1)},{T(i+1,1)}} = 1\) and \(\ind{{U(i,1)},{U(i+1,1)}} = 0\), then
        \begin{equation}
            \maj(T) = \maj(U)
            \quad \text{and} \quad
            \inv(T) = \inv(U) + 1,
        \end{equation}
        \item if \(\ind{{T(i,1)},{T(i+1,1)}} = 0\) and \(\ind{{U(i,1)},{U(i+1,1)}} = 1\), then
        \begin{equation}
            \maj(T) = \maj(U)
            \quad \text{and} \quad
            \inv(T) = \inv(U) - 1,
        \end{equation}
        \item and if \(\ind{{T(i,1)},{T(i+1,1)}} = 0\) and \(\ind{{U(i,1)},{U(i+1,1)}} = 0\), then
        \begin{equation}
            \maj(T) = \maj(U) + (L+1)
            \quad \text{and} \quad
            \inv(T) = \inv(U) - A,
        \end{equation}
    \end{itemize}
    where \(L = \leg(...)\) and \(A = \arm(...)\).
\end{theorem}

\begin{proof}[Proof of \Cref{thm:phi-maj-inv} and \Cref{thm:psi-maj-inv}]
    We prove both theorems simultaneously, by induction on the number of rows in the filling.
    The base case of a single row is straightforward to verify.
    For the inductive step, we use the recursive description of the algorithm.
    The full proof is given in \cite{KLO22}. \todo{This is a placeholder. Write the full proof, that parallels the proof in \cite{KLO22}.}
\end{proof}

\begin{proof}[Alternative proof of \Cref{thm:phi-maj-inv}]
    \todo{Write the proof using the block structure.}
\end{proof}

\begin{theorem} \label{thm:tildeH-rearrangement}
    Let \(\mu\) be a partition and \(\alpha\) be a rearrangement of \(\mu\).
    Then
    \begin{equation}
        \sum_{T \in \Fill[\positives](\mu)} q^{\inv(T)} t^{\maj(T)} \mathbf{x}^T
        =
        \sum_{T \in \Fill[\positives](\alpha)} q^{\inv(T)} t^{\maj(T)} \mathbf{x}^T,
    \end{equation}
    where \(\Fill[\positives](\mu)\) (resp. \(\Fill[\positives](\alpha)\)) is the set of fillings of shape \(\mu\) (resp. \(\alpha\)) with labels in the positive integers \(\positives\).
\end{theorem}

\section{Bijections on augmented fillings and a three-term identity for \texorpdfstring{\(E_\alpha^\sigma\)}{Eασ}} \label{sec:bijective-augmented}

\todo{Write the section on the three-term identity for permuted-basement Macdonald polynomials.}

\begin{theorem}[equivalent to {\cite[Proposition~17]{Ale19}}]\label{thm:Ale10:prop17}
    Let \(\alpha=\alpha_1\cdots\alpha_n\) be a composition such that \(\alpha_i > \alpha_{i+1}\) and \(\sigma\in S_n\) such that \(\sigma_i = \sigma_{i+1} + 1\). Then
    \begin{equation}
        E_{\alpha}^\sigma(X;q;t)
        =
        E_{s_i \permact \alpha}^{\sigma s_i}(X;q;t)
        +
        \frac{1-t}{1-q^{L+1}t^{A}}
        E_{s_i \permact \alpha}^\sigma(X;q;t)
    \end{equation}
    where \(A=\arm(u)\) and \(L=\leg(u)\) for \(u=(i+1,\alpha_{i+1} +1)\in\dg(s_i \permact \alpha)\).
\end{theorem}

We extend this result to any \(\sigma_i>\sigma_{i+1}\).
In the \(q=0\) case, these results are found in \cite[Proposition~5.5]{AS19}.

\begin{theorem}\label{thm:three-term-v1}
For a composition \(\alpha=\alpha_1\cdots\alpha_n\) with \(\alpha_i > \alpha_{i+1}\) and \(\sigma\in S_n\) such that \(\sigma_i>\sigma_{i+1}\), we have
    \begin{equation} \label{eq:three-term-v1}
        E_{\alpha}^\sigma(X;q;t)=E_{s_i \permact \alpha}^{\sigma s_i}(X;q;t)+t^{\rho(\alpha,\sigma,i)}\frac{1-t}{1-q^{L+1}t^{A}}E_{s_i \permact \alpha}^\sigma(X;q;t)
    \end{equation}
    where \(A=\arm(u)\) and \(L=\leg(u)\) for \(u=(i+1,\alpha_{i+1} +1)\in\dg(s_i \permact \alpha)\), and
    \begin{equation}
        \rho(\alpha,\sigma,i)=\twinv(s_i \permact \alpha,\sigma)-\twinv(\alpha,\sigma).
    \end{equation}
\end{theorem}

% \begin{proof}[Algebraic proof]
%     By \cref{thm:Ale}, if $\alpha_i > \alpha_{i+1}$ and $\sigma_i=\sigma_{i+1}+1$, we have
%     \begin{equation}
%         E_{\alpha}^{\sigma}=E_{s_i \permact \alpha}^{\sigma s_i}+\frac{1-t}{1-q^{1+L}t^A}E_{s_i \permact \alpha}^{\sigma}.
%     \end{equation}
%     In particular, for $\sigma=w_0$, we have
%     \begin{align}
%         E_{\alpha}^{w_0}&=E_{s_i \permact \alpha}^{w_0 s_i}+\frac{1-t}{1-q^{1+L}t^A}E_{s_i \permact \alpha}^{w_0}\nonumber\\
%         &=T_{n-i}E_{s_i \permact \alpha}^{w_0}+\frac{1-t}{1-q^{1+L}t^A}E_{s_i \permact \alpha}^{w_0}\label{eq:w0}
%     \end{align}
%     by \eqref{eq:Ti}, since $\twinv(s_i \permact \alpha,w_0s_i)=0$. By \eqref{eq:Ti}, applying $t^{-\twinv(\alpha,\sigma)}T_{\rev(\sigma)}$ to the left hand side gives
%     \begin{equation}
%         t^{-\twinv(\alpha,\sigma)}T_{\rev(\sigma)} E_{\alpha}^{w_0}=E_{\alpha}^\sigma.
%     \end{equation}
%     Now assume $\alpha_i > \alpha_{i+1}$ and $\sigma_i>\sigma_{i+1}$, which implies $\ell(\sigma)>\ell(\sigma s_i)$ and thus $T_{\rev(\sigma s_i)}=T_{\rev(\sigma)}T_{n-i}$. Applying $t^{-\twinv(\alpha,\sigma)}T_{\rev(\sigma)}$ to the right hand side of \eqref{eq:w0} gives
%     \begin{align}
%         E_{\alpha}^\sigma&=t^{-\twinv(\alpha,\sigma)}T_{\rev(\sigma)}T_{n-i}E_{s_i \permact \alpha}^{w_0}
%         +t^{-\twinv(\alpha,\sigma)}\frac{1-t}{1-q^{1+L}t^A}T_{\rev(\sigma)}E_{s_i \permact \alpha}^{w_0}\nonumber\\
%         &=E_{s_i \permact \alpha}^{\sigma s_i}+t^{\twinv(s_i \permact \alpha,\sigma)-\twinv(\alpha,\sigma)}\frac{1-t}{1-q^{1+L}t^A}E_{s_i \permact \alpha}^\sigma \label{eq:f2}
%     \end{align}
% 
%     The final equality \eqref{eq:f2} is obtained by comparing with \eqref{eq:Ti}, using the fact that $\twinv(\alpha,\sigma)=\twinv(s_i \permact \alpha,\sigma s_i)$ when $(\alpha)_i>(\alpha)_{i+1}$ and $\sigma_i>\sigma_{i+1}$.
% \end{proof}

Using the integral forms \(\mathcal{E}_{s_i \permact \alpha}^\sigma(X;q;t)\) instead of the non-integral forms \(E_{s_i \permact \alpha}^\sigma(X;q;t)\), \Cref{eq:three-term-v1} is equivalent to
\begin{equation} \label{eq:three-term-v1-integral}
    \left( 1 - q^{L+1} t^{A + 1} \right) \mathcal{E}_{\alpha}^\sigma
    = 
    \left( 1 - q^{L+1} t^{A} \right) \mathcal{E}_{s_i \permact \alpha}^{\sigma s_i} 
    +
    t^{\rho(\alpha,\sigma,i)} (1 - t) \mathcal{E}_{s_i \permact \alpha}^\sigma.
\end{equation}

% and one over all augmented fillings using the signed alphabet \(\interval{n}_\pm\),
% \begin{equation} \label{eq:calE-Fill-signed}
% 	\mathcal{E}^{\sigma}_{s_i \permact \alpha}(\mathbf{x}; q, t) =
% 	\sum_{T \in \Fill[\interval{n}_\pm](s_i \permact \alpha, \sigma)}
% 	\mathbf{x}^{\content(T)}
% 	q^{\maj(T)} t^{\coinv(T)}
% 	(-t)^{\operatorname{neg}(T)}.
% \end{equation}

Recall that the integral forms have a combinatorial formula in terms of augmented fillings using the signed alphabet \(\interval{n}_\pm\), as given in \Cref{eq:calE-Fill-signed}:
\begin{equation*}
    \mathcal{E}^{\sigma}_{s_i \permact \alpha}(\mathbf{x}; q, t) =
    \sum_{T \in \Fill[\interval{n}_\pm](s_i \permact \alpha, \sigma)}
    \mathbf{x}^{\content(T)}
    q^{\maj(T)} t^{\coinv(T)}
    (-t)^{\operatorname{neg}(T)}.
    \tag{\ref{eq:calE-Fill-signed}}
\end{equation*}

Recall from \Cref{subsec:augmented-to-non-augmented} that any augmented filling \(T \in \Fill(s_i \permact \alpha, \sigma)\) may be identified with a non-augmented filling \(T' \in \Fill(s_i \permact \alpha + \mathbf{1})\) by moving each box one unit up.
This identification preserves \(\content\), \(\maj\), but changes \(\inv\) and \(\coinv\) by a constant depending only on \(s_i \permact \alpha\) and \(\sigma\).
We defined \(\symcoinv(T) = \coinv(T')\) and \(\syminv(T) = \inv(T')\).

We can apply \(\phi\) to \(T'\) to obtain \(U' = \phi(T') \in \Fill(s_i \permact (s_i \permact \alpha + \mathbf{1}))\), and then identify \(U'\) with an augmented filling \(U\) of shape \(\alpha\).
Note that \(\phi\) might or might not change the entries in positions \((i,1)\) and \((i+1,1)\) of \(T'\),
so the basement of \(U\) might be \(\sigma\) or \(\sigma s_i\).

Define the injection \(\phi \colon \Fill(s_i \permact \alpha, \sigma) \to \Fill(\alpha, \sigma) \cup \Fill(\alpha, \sigma s_i)\) by applying the bijection \(\phi\) to the non-augmented filling \(T'\) corresponding to \(T\), and then identifying the result with an augmented filling \(U\).
Similarly, define the injection \(\psi \colon \Fill(s_i \permact \alpha, \sigma) \to \Fill(\alpha, \sigma) \cup \Fill(\alpha, \sigma s_i)\) by applying the bijection \(\psi\) to the non-augmented filling \(T'\) corresponding to \(T\), and then identifying the result with an augmented filling \(U\).

We have immediate coloraries of \Cref{thm:phi-maj-inv} and \Cref{thm:psi-maj-inv}.
\begin{corollary} \label{cor:phi-symcoinv}
    For any \(T \in \Fill(\alpha, \sigma)\) and \(U = \phi(T)\), we have
    \begin{equation}
        \symcoinv(T) = \symcoinv(U)
        \quad \text{and} \quad
        \maj(T) = \maj(U).
    \end{equation}
\end{corollary}

\begin{corollary} \label{cor:psi-symcoinv}
    For any \(T \in \Fill(\alpha, \sigma)\) and \(U = \psi(T) \in \Fill(s_i \permact \alpha, \pi)\), we have
    \begin{itemize}
        \item if \(\ind{{\sigma_{i}},{\sigma_{i+1}}} = 1\) and \(\ind{{\pi_{i}},{\pi_{i+1}}} = 1\), then
        \begin{equation}
            \maj(T) = \maj(U) - (L+1)
            \quad \text{and} \quad
            \symcoinv(T) = \symcoinv(U) - A,
        \end{equation}
        \item if \(\ind{{\sigma_{i}},{\sigma_{i+1}}} = 1\) and \(\ind{{\pi_{i}},{\pi_{i+1}}} = 0\), then
        \begin{equation}
            \maj(T) = \maj(U)
            \quad \text{and} \quad
            \symcoinv(T) = \symcoinv(U) - 1,
        \end{equation}
        \item if \(\ind{{\sigma_{i}},{\sigma_{i+1}}} = 0\) and \(\ind{{\pi_{i}},{\pi_{i+1}}} = 1\), then
        \begin{equation}
            \maj(T) = \maj(U)
            \quad \text{and} \quad
            \symcoinv(T) = \symcoinv(U) + 1,
        \end{equation}
        \item and if \(\ind{{\sigma_{i}},{\sigma_{i+1}}} = 0\) and \(\ind{{\pi_{i}},{\pi_{i+1}}} = 0\), then
        \begin{equation}
            \maj(T) = \maj(U) + (L+1)
            \quad \text{and} \quad
            \symcoinv(T) = \symcoinv(U) + A,
        \end{equation}
    \end{itemize}
    where \(L = \leg(u)\) and \(A = \arm(u)\) for \(u = (i+1, \alpha_{i+1} + 1) \in \dg(s_i \permact \alpha)\).
\end{corollary}

\begin{definition}
    Define \(f \colon \Fill(\alpha, \sigma) \to \Fill(s_i \permact \alpha, \sigma)\)
    and \(g \colon \Fill(\alpha, \sigma) \to \Fill(s_i \permact \alpha, \sigma s_i)\) by
    \begin{equation}
        f(T) =
        \begin{cases}
            \phi(T), & \text{if } \phi(T) \in {\Fill(\alpha, \sigma)}, \\
            \psi(T), & \text{if } \psi(T) \in {\Fill(\alpha, \sigma)},
        \end{cases}
        \quad \text{and} \quad
        g(T) =
        \begin{cases}
            \phi(T), & \text{if } \phi(T) \in {\Fill(\alpha, \sigma s_i)}, \\
            \psi(T), & \text{if } \psi(T) \in {\Fill(\alpha, \sigma s_i)}.
        \end{cases}
    \end{equation}
\end{definition}

% \begin{equation} \label{eq:three-term-a-integral}
%     \left( 1 - q^{L+1} t^{A + 1} \right) \mathcal{E}_{\alpha}^\sigma
%     = 
%     \left( 1 - q^{L+1} t^{A} \right) \mathcal{E}_{s_i \permact \alpha}^{\sigma s_i} 
%     +
%     t^{\rho(\alpha,\sigma,i)} (1 - t) \mathcal{E}_{s_i \permact \alpha}^\sigma
% \end{equation}

For the purposes of this section, let \(\wt(T) = \mathbf{x}^{\content(T)} q^{\maj(T)} t^{\coinv(T)} (-t)^{\operatorname{neg}(T)}\), for an augmented signed filling \(T\).

\begin{lemma} \label{lem:three-term-bijective}
    Assume that \(\alpha_i>\alpha_{i+1}\) and \(\sigma_i > \sigma_{i+1}\).
    Let \(T\) be an augmented filling in \(\Fill(\alpha, \sigma)\).
    Let \(U = f(T)\) and \(V = g(T)\).
    Then
    \begin{equation}
        \left( 1 - q^{L+1} t^{A + 1} \right) \wt(T)
        =
        \left( 1 - q^{L+1} t^{A} \right) \wt(U)
        +
        t^{\rho(\alpha,\sigma,i)} (1 - t) \wt(V).
    \end{equation}
\end{lemma}

\begin{proof}
    If \(U = \phi(T)\) and \(V = \psi(T)\),
    then \(\wt(U) = \wt(T)\) by \Cref{cor:phi-symcoinv},
    and \(\wt(V) = t^{-\rho(\alpha,\sigma,i)} q^{L+1} t^{A} \wt(T)\) by \Cref{cor:psi-symcoinv}.
    Thus,
    \begin{multline}
        \left( 1 - q^{L+1} t^{A} \right) \wt(U)
        +
        t^{\rho(\alpha,\sigma,i)} (1 - t) \wt(V)
        = \\
        \begin{aligned}
        &=
        \left( 1 - q^{L+1} t^{A} \right) \wt(T)
        +
        t^{\rho(\alpha,\sigma,i)} (1 - t) t^{-\rho(\alpha,\sigma,i)} q^{L+1} t^{A} \wt(T) \\
        &=
        \left( 1 - q^{L+1} t^{A} + (1 - t) q^{L+1} t^{A + 1} \right) \wt(T) \\
        &=
        \left( 1 - q^{L+1} t^{A + 1} \right) \wt(T).
        \end{aligned}
    \end{multline}
    If \(U = \psi(T)\) and \(V = \phi(T)\),
    then \(\wt(U) = t \wt(T)\) by \Cref{cor:psi-symcoinv},
    and \(\wt(V) = t^{-\rho(\alpha,\sigma,i)} \wt(T)\) by \Cref{cor:phi-symcoinv}.
    Thus,
    \begin{multline}
        \left( 1 - q^{L+1} t^{A} \right) \wt(U)
        +
        t^{\rho(\alpha,\sigma,i)} (1 - t) \wt(V)
        = \\
        \begin{aligned}
        &=
        \left( 1 - q^{L+1} t^{A} \right) t \wt(T)
        +
        t^{\rho(\alpha,\sigma,i)} (1 - t) t^{-\rho(\alpha,\sigma,i)} \wt(T) \\
        &=
        \left( t - q^{L+1} t^{A + 1} + 1 - t \right) \wt(T) \\
        &=
        \left( 1 - q^{L+1} t^{A + 1} \right) \wt(T).
        \end{aligned}
    \end{multline}
\end{proof}

\begin{corollary} \label{cor:three-term-a}
    Let \(\alpha\) be a composition such that \(\alpha_i > \alpha_{i+1}\) and \(\sigma \in S_n\) such that \(\sigma_i > \sigma_{i+1}\).
    Then,
    \begin{equation}
        \left( 1 - q^{L+1} t^{A + 1} \right) \mathcal{E}_{\alpha}^\sigma
        = 
        \left( 1 - q^{L+1} t^{A} \right) \mathcal{E}_{s_i \permact \alpha}^{\sigma s_i} 
        +
        t^{\rho(\alpha,\sigma,i)} (1 - t) \mathcal{E}_{s_i \permact \alpha}^\sigma,
    \end{equation}
    where \(A=\arm(u)\) and \(L=\leg(u)\) for \(u=(i+1,\alpha_{i+1} +1)\in\dg(s_i \permact \alpha)\), and \( \rho(\alpha,\sigma,i)=\twinv(s_i \permact \alpha,\sigma)-\twinv(\alpha,\sigma)\).
\end{corollary}

\begin{proof}
    This follows by summing the equation in \Cref{lem:three-term-bijective} over all \(T \in \Fill(\alpha, \sigma)\).
\end{proof}

\begin{theorem} \label{thm:three-term-a-non-integral}
    Let \(\alpha\) be a composition such that \(\alpha_i > \alpha_{i+1}\) and \(\sigma \in S_n\) such that \(\sigma_i > \sigma_{i+1}\).
    Then,
    \begin{equation}
        E_{\alpha}^\sigma(X;q;t)=E_{s_i \permact \alpha}^{\sigma s_i}(X;q;t)+t^{\rho(\alpha,\sigma,i)}\frac{1-t}{1-q^{L+1}t^{A}}E_{s_i \permact \alpha}^\sigma(X;q;t)
    \end{equation}
    where \(A=\arm(u)\) and \(L=\leg(u)\) for \(u=(i+1,\alpha_{i+1} +1)\in\dg(s_i \permact \alpha)\), and  \( \rho(\alpha,\sigma,i)=\twinv(s_i \permact \alpha,\sigma)-\twinv(\alpha,\sigma)\).
\end{theorem}

\begin{proof}
    This follows from \Cref{cor:three-term-bijective} by dividing both sides of \Cref{eq:three-term-v1-integral} by the appropriate factors to convert the integral forms to non-integral forms.
\end{proof}

% -------

\begin{corollary} \label{cor:three-term-b}
    Let \(i \in \interval{n-1}\),
    \(\alpha = \composition{\alpha_1, \dots, \alpha_n}\) be a composition with \(\alpha_i < \alpha_{i+1}\),
    and \(\sigma = \window{\sigma_1, \dots, \sigma_n} \in \sym{n}\) be a permutation with \(\sigma_i < \sigma_{i+1}\).
    Then, we have
    \begin{equation}
        \left( 1 - q^{L+1} t^{A + 1} \right) \mathcal{E}_{s_i \permact \alpha}^\sigma
        = 
        \left( 1 - q^{L+1} t^{A} \right) \mathcal{E}_{\alpha}^{\sigma} 
        +
        (q^{L+1} t^A) (1 - t) t^{\twinv(\alpha, \sigma) - \twinv(s_i \permact \alpha, \sigma)}
        \mathcal{E}_{\alpha}^{s_i \sigma},
    \end{equation}
    where \(A=\arm(u)\) and \(L=\leg(u)\) for \(u=(i+1,\alpha_i +1)\in\dg(\alpha)\).
\end{corollary}

\begin{theorem} \label{thm:three-term-b-non-integral}
    Let \(i \in \interval{n-1}\),
    \(\alpha = \composition{\alpha_1, \dots, \alpha_n}\) be a composition with \(\alpha_i < \alpha_{i+1}\),
    and \(\sigma = \window{\sigma_1, \dots, \sigma_n} \in \sym{n}\) be a permutation with \(\sigma_i < \sigma_{i+1}\).
    Then, we have
    \begin{equation} \label{eq:three-term-b}
        E_{s_i \permact \alpha}^\sigma(\mathbf{x}; q, t)
        =
        E_\alpha^\sigma(\mathbf{x}; q, t)
        +
        (q^{L+1} t^A)
        t^{\twinv(\alpha, \sigma) - \twinv(s_i \permact \alpha, \sigma)}
        \frac{1 - t}{1 - q^{L+1} t^{A}}
        E_\alpha^{s_i \sigma}(\mathbf{x}; q, t),
    \end{equation}
    where \(L = \leg(x)\) and \(A = \arm(x)\) for \(x = (i+1, \alpha_i + 1) \in \dg(\alpha)\).
\end{theorem}