\chapter{Probabilistic bijective combinatorics for Macdonald polynomials} \label{chap:probabilistic}

\textcite[Theorem~22]{Ale19} established a partial symmetry result
for permuted-basement Macdonald polynomials under certain conditions.
Using \cite[Lemma~20 and Proposition~15]{Ale19},
we reformulate the result as follows.

\begin{restatable}[equivalent to {\cite[Theorem~22]{Ale19}}]{theorem}{symmetryAle}
	\label{theorem:symmetry-Ale19}
	Let \(\alpha=\composition{\alpha_1,\ldots,\alpha_n}\) be a composition
	with \(\alpha_i=\alpha_{i+1}\) for some \(i\in\interval{n-1}\).
	Let \(\sigma \in \sym{n}\), such that \(\sigma_{i+1} = \sigma_i \pm 1\).
	Then,
	\begin{equation*}
		E_{\alpha}^{\sigma}(\mathbf{x}; q, t)
		= E_{\alpha}^{\sigma s_i}(\mathbf{x}; q, t).
	\end{equation*}
\end{restatable}

The main result of this chapter is stated in the following theorem,
which extends \cref{theorem:symmetry-Ale19}
by removing the assumption that \(\sigma_i\) and \(\sigma_{i+1}\) are consecutive.

\begin{restatable}{theorem}{main}
	\label{theorem:main}
	Let \(\alpha=\composition{\alpha_1,\ldots,\alpha_n}\) be a composition
	with \(\alpha_i=\alpha_{i+1}\) for some \(i\in\interval{n-1}\),
	and let \(\sigma \in \sym{n}\).
	Then,
	\begin{equation*}
		E_{\alpha}^{\sigma}(\mathbf{x}; q, t)
		= E_{\alpha}^{\sigma s_i}(\mathbf{x}; q, t).
	\end{equation*}
\end{restatable}

The combinatorial interpretation of \cref{theorem:main} is that
swapping the basement entries in two adjacent columns of the same height
does not change the generating function for non-attacking fillings.
Our approach for proving this builds upon a (classical) bijection
introduced in \cite{CHMMW22} for unrestricted fillings of partition shape,
in which two adjacent entries of the filling
in the bottom row of columns of the same height are swapped
while preserving the overall weight of the filling.
\cite{Man24} extended this technique
by introducing a probabilistic bijection
for partition-shaped non-attacking fillings.
In this chapter,
we generalize this probabilistic bijection
to composition-shaped non-attacking fillings,
thereby giving a bijective proof of \cref{theorem:main}.

As implications of \cref{theorem:main},
certain assumptions can be removed from
theorems in \cite{Ale19} and \cite{CMW22},
which we summarize in \cref{sec:applications}.

\section{Probabilistic bijections}

Probabilistic bijections,
also referred to as \emph{bijectivization} or \emph{coupling} by \cite{BP19},
generalize weight-preserving bijections.
We follow the exposition given in~\cite{AF22}.

\begin{definition}[Probability map]
	Let \(\algebra\) be an algebra.
	Let \(\mathbf{T}\) and \(\mathbf{U}\) be sets.
	A \vocab{probability map} from \(\mathbf{T}\) to \(\mathbf{U}\) is
	a function \(\prob{} \colon \mathbf{T} \times \mathbf{U} \to \algebra\) such that
	\begin{equation}
		\sum_{U \in\mathbf{U}} \prob{}(T, U) = 1
	\end{equation}
	for every \(T \in \mathbf{T}\).
\end{definition}

Note that these do not formally correspond to probability maps
in the sense of probability theory
because the values of \(\prob{}\)
are not required to be in the interval \([0, 1]\) or even real numbers.

\begin{definition}[Probabilistic bijection {\cite{AF22}}]\label{def:prob}
	Let \(\algebra\) be an algebra.
	Let \(\mathbf{T}\) and \(\mathbf{U}\) be finite sets
	equipped with weight functions
	\(\wt_{\mathbf{T}} \colon \mathbf{T} \to \algebra\) and
	\(\wt_{\mathbf{U}} \colon \mathbf{U} \to \algebra\).
	A \vocab{probabilistic bijection} between
	\( (\mathbf{T}, \wt_{\mathbf{T}}) \) and
	\( (\mathbf{U}, \wt_{\mathbf{U}}) \) is
	a pair of probability maps
	\(\prob{\mathbf{T}} \colon \mathbf{T} \times \mathbf{U} \to \algebra\) and
	\(\prob{\mathbf{U}} \colon \mathbf{U} \times \mathbf{T} \to \algebra\)
	such that
	\begin{equation}
		\wt_{\mathbf{T}}(T) \prob{\mathbf{T}}(T, U)
		= \wt_{\mathbf{U}}(U) \prob{\mathbf{U}}(U, T)
	\end{equation}
	for every \(T \in \mathbf{T}\) and \(U \in \mathbf{U}\).
\end{definition}

A weight-preserving bijection \( f \colon \mathbf{T} \to \mathbf{U} \)
canonically induces a probabilistic bijection
between \((\mathbf{T}, \wt_{\mathbf{T}})\) and \((\mathbf{U}, \wt_{\mathbf{U}})\)
by setting
\begin{equation}
	\prob{\mathbf{T}}(T, U) = \prob{\mathbf{U}}(U, T) =
	\begin{cases}
		1 & \text{if } f(T) = U, \\
		0 & \text{otherwise},
	\end{cases}
\end{equation}
explaining the choice of terminology.
Moreover, as shown in \cref{proposition:equal-weight-sum},
this generalization retains a key property of weight-preserving bijections:
the sum of the weights of \(\mathbf{T}\) and of \(\mathbf{U}\) are equal.

\begin{proposition}
	\label{proposition:equal-weight-sum}
	Let \(\prob{\mathbf{T}} \colon \mathbf{T} \times \mathbf{U} \to \algebra\)
	and \(\prob{\mathbf{U}} \colon \mathbf{U} \times \mathbf{T} \to \algebra\)
	form a probabilistic bijection between
	\( (\mathbf{T}, \wt_{\mathbf{T}}) \) and
	\( (\mathbf{U}, \wt_{\mathbf{U}}) \).
	Then,
	\begin{equation}
		\sum_{T \in \mathbf{T}} \wt(T) = \sum_{U \in\mathbf{U}} \wt(U).
	\end{equation}
\end{proposition}

\begin{proof}
	We compute
	\begin{align}
		\sum_{T \in \mathbf{T}} \wt_{\mathbf{T}}(T)
		&= \sum_{T \in \mathbf{T}} \sum_{U \in \mathbf{U}}
			\wt_{\mathbf{T}}(T) \prob{\mathbf{T}}(T, U) \\
		&= \sum_{U \in \mathbf{U}} \sum_{T \in \mathbf{T}}
			\wt_{\mathbf{U}}(U) \prob{\mathbf{U}}(U, T)
		= \sum_{U \in \mathbf{U}} \wt_{\mathbf{U}}(U). \qedhere
	\end{align}
\end{proof}

A Markov-theoretical interpretation of a probabilistic bijection is that
it defines a bipartite Markov chain on the state space \( \mathbf{T} \cup \mathbf{U} \)
with transition probabilities given by the function \( \prob{} \) under a specific balance condition.
In the case of \cref{def:prob}, this condition is detailed balance,
but a probabilistic bijection can also be defined using a more general balance condition.

\section{A probabilistic bijection for non-attacking fillings}
\label{sec:probabilistic}

In this section,
we define a map on non-attacking fillings
that allows us to construct a probabilistic bijection between
\(\NAF(\alpha,\sigma)\) and \(\NAF(\alpha,\sigma s_i)\)
in order to prove \cref{theorem:main} in \cref{sec:proof}.

In \cite{CHMMW22},
a weight-preserving bijection was established between
fillings in \(\Fill(\lambda)\) with statistics \(\maj\) and \(\inv\),
corresponding to swapping two adjacent entries in the bottom row within columns of the same height.
This was done using an operator \(\tau_i\),
originally introduced in \cite{LN12},
on fillings in \(\Fill(\lambda)\).
The bijection established the equality of generating functions
over weighted fillings for the sets
\(\{T\in\Fill(\lambda) \suchthat \text{bottom row of \(T\) is \(w\)}\}\)
and
\(\{T\in\Fill(\lambda) \suchthat \text{bottom row of \(T\) is \(w s_i\)}\}\)
for an index \(i\) such that \(\lambda_i = \lambda_{i+1}\).
The result was used to obtain a compact formula for the modified Macdonald polynomials.

In our case,
proving that \(E_\alpha^\sigma=E_\alpha^{\sigma s_i}\) combinatorially
requires showing that the weight generating functions
over \(\NAF(\alpha,\sigma)\) and \(\NAF(\alpha,\sigma s_i)\) are equal,
which also amounts to finding a bijection on tableaux
that swaps basement entries in columns \(i\) and \(i+1\).
However,
the operator \(\tau_i\) used in \cite{CHMMW22}
does not preserve the non-attacking property of tableaux,
so the same strategy could not be applied.
This issue was resolved in \cite{Man24}
by introducing a probabilistic operator \(\widehat{\tau}_i\)
for non-attacking tableaux with partition shape,
in order to obtain a compact formula for \(P_{\lambda}\).
The latter operator is the map we use to define our probabilistic bijection.

There are two key differences between our setting and the one in \cite{Man24}.
First, our map applies to tableaux with composition shapes,
whereas \cite{Man24} considers only partition shapes.
This distinction allows \cite{Man24} to restrict attention to type A triples,
while we consider both types.
Second, our fillings include basements,
eliminating the need for the so-called \emph{degenerate} triples in \cite{Man24}.
Hence, the probability map in \cref{definition:prob}
and the surrounding lemmas are adapted from the original to fit our setting.

Throughout this section,
let \(\alpha=\composition{\alpha_1,\ldots,\alpha_n}\) be a composition,
let \(\sigma \in \sym{n}\) be a permutation,
and let \(i \in \interval{n-1}\) such that \(\alpha_i = \alpha_{i+1}\).

\begin{definition}[{\(\swap[r]{i}\) and \(\sswap[0, r]{i}\)}]
	\label{definition:swap}
	Given a filling \(T\) of shape \(\alpha\) and \(r \in \interval{0, \alpha_i}\),
	let \(\swap[r]{i}(T)\) denote the filling obtained by
	swapping the entries in boxes \((i, r)\) and \((i+1, r)\) in \(T\).
	Moreover,
	let \(\sswap[0, r]{i} = \swap[r]{i} \circ \cdots \circ \swap[1]{i} \circ \swap[0]{i}\),
	which represents a sequence of swaps
	in the first \(r+1\) rows within the columns \(i\) and \(i+1\).
\end{definition}

Note that if \(T \in \Fill(\alpha, \sigma)\),
then \(\sswap[0, r]{i}(T) \in \Fill(\alpha, \sigma s_i)\) for any \(r\geq 0\).
Also, the fillings \(T\) and \(\sswap[0, r]{i}(T)\) have the same content.
However, \(\swap[r]{i}\) and \(\sswap[0, r]{i}\)
do not in general preserve the property of a filling being non-attacking.

\begin{definition}[{\(\pp[r]{i}\)}]
	\label{definition:propagate}
	Let \(T \in \NAF(\alpha, \sigma)\) be a non-attacking filling.
	Let \(r \in \interval{0, \alpha_i-1}\).
	We define \(\pp[r]{i}(T) \in \rationals(q, t)\).
	Let \(a = T(i, r)\), \(b = T(i+1, r)\), \(c = T(i, r+1)\), and \(d = T(i+1, r+1)\).
	Let \(A = \arm(i+1, r+1)\) and \(\ell = \leg(i+1, r+1)\).

	If \(|\{a, b, c, d\}| = 4\), then we split into two cases.
	\begin{enumerate}
		\item \label{item:pp-cda-cdb}
			If \(\ind{c, d, a} = \ind{c, d, b}\), then \(\pp[r]{i}(T) = 0\).
		\item \label{item:pp-cda-dcb}
			If \(\ind{c, d, a} = \ind{d, c, b}\), then \(\pp[r]{i}(T) = 1\).
	\end{enumerate}

	If \(|\{a, b, c, d\}| = 3\),
	the non-attacking condition implies that \(a \neq b\),
	\(a \neq d\) and \(c \neq d\),
	so we have three cases.
	\begin{enumerate}[resume]
		\item \label{item:pp-bc}
			If \(b = c\), then \(\pp[r]{i}(T) = 0\).
		\item \label{item:pp-bd}
			If \(b = d\), then \(\pp[r]{i}(T) = 1\).
		\item \label{item:pp-ac}
			If \(a = c\), then
			\begin{equation*}
				\pp[r]{i}(T) =
				t^{1 - \ind{d, a, b}}
				\frac{1 - q^{\ell+1}t^{A+1}}{1-q^{\ell+1}t^{A+2}}.
			\end{equation*}
	\end{enumerate}

	If \(|\{a, b, c, d\}| = 2\),
	then we have one case.
	\begin{enumerate}[resume]
		\item \label{item:pp-ac-bd}
		If \(a = c\) and \(b = d\), then \(\pp[r]{i}(T) = 1\).
	\end{enumerate}
\end{definition}

\Cref{table:propagate} summarizes the values of \(\pp[r]{i}(T)\)
for \(T \in \NAF(\alpha, \sigma)\) and \(r \in \interval{0, \alpha_i-1}\).
\begin{table}[htbp]
	\centering
	\ytableausetup{boxsize=1em}
	\setcellgapes{0pt}
	\makegapedcells
	\begin{tabular}[c]{rccc}
		\toprule
		& \(T\) & \(\pp[r]{i}(T)\) & \(1 - \pp[r]{i}(T)\) \\
		\midrule
		\ref{item:pp-cda-cdb}
		&
		\begin{tabular}[t]{c}
		\begin{ytableau}
			\none[\scriptstyle r+1] & \none & c & d \\
			\none[\scriptstyle r]   & \none & a & b \\
		\end{ytableau} \\
		{\scriptsize \(\ind{c, d, a} = \ind{c, d, b}\)}
		\end{tabular}
		& \(0\) & \(1\) \\ \midrule
		\ref{item:pp-cda-dcb}
		&
		\begin{tabular}[t]{c}
		\begin{ytableau}
			\none[\scriptstyle r+1] & \none & c & d \\
			\none[\scriptstyle r]   & \none & a & b \\
		\end{ytableau}\\
		{\scriptsize \(\ind{c, d, a} = \ind{d, c, b}\)}
		\end{tabular}
		& \(1\) & \(0\) \\ \midrule
		\ref{item:pp-bc}
		& \begin{ytableau}
				\none[\scriptstyle r+1] & \none & b & d \\
				\none[\scriptstyle r]   & \none & a & b \\
			\end{ytableau}
		& \(0\) & \(1\) \\ \midrule
		\ref{item:pp-bd}
		& \begin{ytableau}
				\none[\scriptstyle r+1] & \none & c & b \\
				\none[\scriptstyle r]   & \none & a & b \\
			\end{ytableau}
		& \(1\) & \(0\) \\ \midrule
		\ref{item:pp-ac}
		& \begin{ytableau}
				\none[\scriptstyle r+1] & \none & a & d \\
				\none[\scriptstyle r]   & \none & a & b \\
			\end{ytableau}
		& \(t^{1 - \ind{d, a, b}} \frac{1 - q^{\ell+1}t^{A+1}}{1-q^{\ell+1}t^{A+2}}\)
		& \((q^{\ell+1}t^{A+1})^{\ind{d, a, b}} \frac{1-t}{1 - q^{\ell+1}t^{A+2}}\) \\ \midrule
		\ref{item:pp-ac-bd}
		& \begin{ytableau}
				\none[\scriptstyle r+1] & \none & a & b \\
				\none[\scriptstyle r]   & \none & a & b \\
				\none & \none[\scriptstyle i] & \none[\scriptstyle i+1]
			\end{ytableau}
		& \(1\) & \(0\) \\
		\bottomrule
	\end{tabular}
	\caption{
		The values of \(\pp[r]{i}(T)\) and \(1 - \pp[r]{i}(T)\)
		for \(T \in \NAF(\alpha, \sigma)\)
		and \(r \in \interval{0, \alpha_i-1}\),
		on each case of \cref{definition:propagate}.
	}
	\label{table:propagate}
\end{table}

For ease of computation,
we note that, in case \ref{item:pp-ac},
we have
\begin{equation*}
	1 - \pp[r]{i}(T)
	=
	(q^{\ell+1}t^{A+1})^{\ind{d, a, b}}
	\frac{1-t}{1 - q^{\ell+1}t^{A+2}},
\end{equation*}
which is straightforward to verify by
splitting into cases of \(\ind{d, a, b}\) being \(0\) or \(1\)
(compare to {\cite[Lemma 3.3]{Man24}}).

As a convention, we set \(\pp[\alpha_i]{i}(T) = 0\).

\begin{definition}[\(\prob{i}\)] \label{definition:prob}
	The map
	\begin{equation*}
		\prob{i} \colon
		\NAF(\alpha, \sigma) \times \Fill(\alpha, \sigma s_i)
		\to \rationals(q, t)
	\end{equation*}
	is defined for \(T \in \NAF(\alpha, \sigma)\) and
	\(U \in \Fill(\alpha, \sigma s_i)\) by letting
	\begin{equation}\label{eq:prob factors}
		\prob{i}(T, U) =
		\left( \prod_{r = 0}^{h-1} \pp[r]{i}(T) \right)
		\left( 1 - \pp[h]{i}(T) \right)
	\end{equation}
	if there exists \(h \in \interval{0, \alpha_i}\)
	such that \(U = \sswap[0, h]{i}(T)\),
	and \(\prob{i}(T, U) = 0\) otherwise.
\end{definition}

\begin{proposition}
	The map
	\(\prob{i} \colon \NAF(\alpha, \sigma) \times \Fill(\alpha, \sigma s_i) \to \rationals(q, t)\)
	is a probability map, that is,
	for all \(T \in \NAF(\alpha, \sigma)\),
	\begin{equation*}
		\sum_{U \in \Fill(\alpha, \sigma s_i)} \prob{i}(T, U) = 1.
	\end{equation*}
\end{proposition}

\begin{proof}
	Using that \(\pp[\alpha_i]{i}(T) = 0\),
	it is straightforward to compute
	\begin{equation*}
		\sum_{h=0}^{\alpha_i}
		\left( \prod_{r = 0}^{h-1} \pp[r]{i}(T) \right)
		\left( 1 - \pp[h]{i}(T) \right)
		= 1,
	\end{equation*}
	as required.
\end{proof}

\begin{example}
	Let \(T \in \NAF(\composition{3, 4, 4, 0, 0}, \window{5, 1, 3, 4, 2})\)
	be the non-attacking filling
	\begin{equation*}
		\ytableausetup{aligntableaux=center}
		\begin{ytableau}
			\none & 3 & 2 \\
			4 & 1 & 2 \\
			3 & 4 & 1 \\
			3 & 4 & 2 \\
			5 & 1 & 3 & 4 & 2
		\end{ytableau}.
	\end{equation*}
	The values of \(\pp[r]{2}(T)\) and \(1 - \pp[r]{2}(T)\) are
	\begin{equation*}
		\begin{aligned}
			\pp[0]{2}(T) &= 1,
			& 1-\pp[0]{2}(T) &= 0, \\
			\pp[1]{2}(T) &= \frac{1 - q^{3}t^{1}}{1 - q^{3}t^{2}},
			& 1-\pp[1]{2}(T) &= \frac{q^{3}t^{1} - q^{3}t^{2}}{1 - q^{3}t^{2}}, \\
			\pp[2]{2}(T) &= 0,
			& 1-\pp[2]{2}(T) &= 1, \\
			\pp[3]{2}(T) &= 1,
			& 1-\pp[3]{2}(T) &= 0, \\
			\pp[4]{2}(T) &= 0,
			& 1-\pp[4]{2}(T) &= 1.
		\end{aligned}
	\end{equation*}
	Hence, the values of \(\prob{2}{T, \sswap[0, h]{2}(T)}\) for \(h \in \interval{0, 4}\) are
	\begin{equation*}
		\begin{aligned}
			\prob{2}\left(T, {\sswap[0, 0]{2}(T)}\right) &= 0, \\
			\prob{2}\left(T, {\sswap[0, 1]{2}(T)}\right) &= \frac{q^{3}t^{1} - q^{3}t^{2}}{1 - q^{3}t^{2}}, \\
			\prob{2}\left(T, {\sswap[0, 2]{2}(T)}\right) &= \frac{1 - q^{3}t^{1}}{1 - q^{3}t^{2}}, \\
			\prob{2}\left(T, {\sswap[0, 3]{2}(T)}\right) &= 0, \\
			\prob{2}\left(T, {\sswap[0, 4]{2}(T)}\right) &= 0.
		\end{aligned}
	\end{equation*}
	The two nonzero values of \(\prob{2}(T, U)\) are obtained for
	\(U = \sswap[0, 1]{2}(T)\) and
	\(U = \sswap[0, 2]{2}(T)\),
	which are, respectively,
	\begin{equation*}
		\ytableausetup{aligntableaux=center}
		\begin{ytableau}
			\none & 3 & 2 \\
			4 & 1 & 2 \\
			3 & 4 & 1 \\
			3 & 2 & 4 \\
			5 & 3 & 1 & 4 & 2
		\end{ytableau}
		\quad \text{and} \qquad
		\begin{ytableau}
			\none & 3 & 2 \\
			4 & 1 & 2 \\
			3 & 1 & 4 \\
			3 & 2 & 4 \\
			5 & 3 & 1 & 4 & 2
		\end{ytableau}.
	\end{equation*}
\end{example}

By construction, the \(\prob{i}\) is nonzero only for pairs of non-attacking fillings.
\begin{proposition} \label{proposition:restrict-to-naf-naf}
	Let \(T \in \NAF(\alpha, \sigma)\) be a non-attacking filling.
	Let \(U \in \Fill(\alpha, \sigma s_i)\) be a filling.
	If \(\prob{i}(T, U) \neq 0\), then \(U\) is non-attacking.
\end{proposition}

\begin{proof}
	Note that attacking pairs not involving columns \(i\) and \(i+1\)
	remain unchanged between \(T\) and \(U\),
	and attacking pairs involving one box in columns \(i\) or \(i+1\) in \(T\)
	correspond to attacking pairs with a box in the same columns in \(U\),
	though their positions may be swapped between the two columns.
	Hence,
	the only distinction in attacking pairs between \(T\) and \(U\)
	comes from the attacking boxes in both columns \(i\) and \(i+1\).

	If \(\prob{i}(T, U) \neq 0\),
	there exists \(h \in \interval{0, \alpha_i}\) such that \(U = \sswap[0, h]{i}(T)\),
	and thus by \Cref{eq:prob factors},
	\(\pp[r]{i}(T) \neq 0\) for all \(r \in \interval{0, h-1}\) and \(\pp[h]{i}(T) \neq 1\).
	Therefore,
	it suffices to show that if \(T\) is non-attacking,
	then \(U(i, r) \neq U(i+1, r+1)\) for all \(r \in \interval{0, h}\).
	For \(r \in \interval{0, h-1}\),
	we have \(\pp[r]{i}(T) \neq 0\), so the configuration of \cref{item:pp-bc} does not occur;
	thus, \(U(i, r) = T(i+1, r) \neq T(i, r+1) = U(i+1, r+1)\).
	For \(r = h\),
	we have \(\pp[h]{i}(T) \neq 1\),
	so configurations of \cref{item:pp-bd} or \cref{item:pp-ac-bd} does not occur;
	thus, \(U(i, h) = T(i+1, h) \neq T(i+1, h+1) = U(i+1, h+1)\).
	Therefore, \(U\) is indeed non-attacking.
\end{proof}

\begin{corollary}
	The restriction
	\begin{equation*}
		\prob{i} \colon \NAF(\alpha, \sigma) \times \NAF(\alpha, \sigma s_i) \to \rationals(q, t)
	\end{equation*}
	is a probability map.
\end{corollary}

As an abuse of notation, we write \(\prob{i}\) for both
the function on \(\NAF(\alpha, \sigma) \times \Fill(\alpha, \sigma s_i)\) and
the function on \(\NAF(\alpha, \sigma s_i) \times \NAF(\alpha, \sigma)\).
Moreover, we also write \(\prob{i}\) for both
the function on \(\NAF(\alpha, \sigma) \times \NAF(\alpha, \sigma s_i)\) and
the function on \(\NAF(\alpha, \sigma s_i) \times \NAF(\alpha, \sigma)\).

The main result of this section is \cref{theorem:wt-prob},
which generalizes \cite[Lemma 4.4]{Man24} to our setting.

\begin{theorem}[adapted from {\cite[Lemma 4.4]{Man24}}]
	\label{theorem:wt-prob}
	The pair of maps
	%\begin{gather*}
	%	{\prob{i}} \colon \NAF(\alpha, \sigma) \times \NAF(\alpha, \sigma s_i) \to \rationals(q, t) \\
	%	{\prob{i}} \colon \NAF(\alpha, \sigma s_i) \times \NAF(\alpha, \sigma) \to \rationals(q, t)
	%\end{gather*}
	defines a probabilistic bijection
	between \(\NAF(\alpha, \sigma)\) and \(\NAF(\alpha, \sigma s_i)\)
	with respect to the weight function \(\wtqt\).
	In other words,
	for all \(T \in \NAF(\alpha, \sigma)\)
	and \(U \in \NAF(\alpha, \sigma s_i)\),
	\begin{equation*}
		\wtqt(U)\prob{i}(U, T) = \wtqt(T)\prob{i}(T, U).
	\end{equation*}
\end{theorem}

For organization purposes,
we split the proof of \cref{theorem:wt-prob} into several lemmas.

\begin{lemma} \label{lemma:useful-ind-five-variables}
	Let \(a, b, c, d, f \in [n]\).
	Then,
	\begin{equation*}
		\ind{c, f, a} + \ind{d, f, b} - \ind{c, f, b} - \ind{d, f, a}
		= \ind{c, b} + \ind{d, a} - \ind{c, a} - \ind{d, b}.
	\end{equation*}
\end{lemma}
\begin{proof}
	We directly compute
	\begin{align*}
		\ind{c, f, a} + \ind{d, f, b} - \ind{c, f, b} - \ind{d, f, a}
		&= (\ind{c, f} + \ind{f, a} - \ind{c, a}) + (\ind{d, f} + \ind{f, b} - \ind{d, b}) \\
		&\mathrel{\vphantom{=}} - (\ind{c, f} + \ind{f, b} - \ind{c, b}) - (\ind{d, f} + \ind{f, a} - \ind{d, a}) \\
		&= \ind{c, b} + \ind{d, a} - \ind{c, a} - \ind{d, b}. \qedhere
	\end{align*}
\end{proof}

The \(q,t\)-weight of a filling can be decomposed as the product of the weights contributed by the descents within each row and the inversions between every pair of rows.
Let \(L = \max_i \alpha_i\) be the total number of rows.
For \(0\leq r\leq L-1\),
define \(\dg[r]{(\alpha)}\) as
the set of boxes in the \(r\)\textsuperscript{th} row of \(\dg(\alpha)\).
Then, define
\begin{equation*}
	\maj[r + 1](T) = \sum_{u \in \Des[r + 1](T)} \leg(u) + 1,
\end{equation*}
where \(\Des[r+1](T)=\Des(T)\cap \dg[r+1]{(\alpha)}\).
Also, define \(\coinv[r, r+1](T)\) as
the number of inversion triples in \(T\)
within the pair of rows \(r\) and \(r + 1\).
Then, the \(r\)\textsuperscript{th} component
of the \(q, t\)-weight
of a non-attacking filling \(T \in \NAF(\alpha, \sigma)\)
is defined as
\begin{equation}
	\label{eq:rthcomponent}
	\wtqt[r](T) =
	q^{\maj[r + 1](T)}
	t^{\coinv[r, r + 1](T)}
	\prod_{\substack{u \in \dg[r + 1]{(\alpha)} \\ T(u) \neq T(\south{u})}}
	\frac{1 - t}{1 - q^{1 + \leg(u)} t^{1 + \arm(u)}},
\end{equation}
yielding the decomposition
\begin{equation*}
	\wtqt(T) = \prod_{r = 0}^{L-1} \wtqt[r](T).
\end{equation*}

\begin{lemma} \label{lemma:wtqt-prob-balanced-r}
	Let \(T \in \NAF(\alpha, \sigma)\) be a non-attacking filling.
	Let \(U = \sswap[0, h]{i}(T)\) for some \(h \in \interval{0, \alpha_i}\).
	Then, for all \(r \in \interval{0, h - 1}\),
	\begin{equation} \label{eq:wtqt-prob-balanced-r}
		\wtqt[r](U)\pp[r]{i}(U) = \wtqt[r](T)\pp[r]{i}(T).
	\end{equation}
\end{lemma}

\begin{proof}
	Let \(r \in \interval{0, h-1}\).
	Let
	\begin{align*}
		c & = T(i, r+1)=U(i+1, r+1), &
		d & = T(i+1, r+1)=U(i, r+1), \\
		a & = T(i, r)=U(i+1, r), \text{ and } &
		b & = T(i+1, r)=U(i, r).
	\end{align*}
	Since \(U\) is non-attacking, we have \(b \neq c\).
	We note that \(\maj[r + 1](T) = \maj[r + 1](U)\).
	Moreover,
	triples not involving columns \(i\) and \(i+1\)
	are preserved between \(T\) and \(U\),
	triples involving column \(i\) in \(T\)
	correspond to triples involving column \(i+1\) in \(U\),
	and triples involving column \(i+1\) in \(T\)
	correspond to triples involving column \(i\) in \(U\).
	Hence, the only distinction in the coinversion and inversion statistics between \(T\) and \(U\)
	comes from the triples involving both columns \(i\) and \(i+1\).
	Therefore,
	\begin{equation*}
		\coinv[r, r + 1](T) - \coinv[r, r + 1](U)
		=
		\inv[r, r + 1](U) - \inv[r, r + 1](T)
		=
		\ind{d, c, b} - \ind{c, d, a}.
	\end{equation*}
	We show that \Cref{eq:wtqt-prob-balanced-r} holds
	for all cases of \((a,b,c,d)\) with \(b\neq c\).

	First, suppose \(|\{a, b, c, d\}| = 4\).
	Then,
	the sets \(\{u \in \Arm(i+1, r+1) \suchthat T(u) \neq T(\south{u})\}\)
	and \(\{u \in \Arm(i+1, r+1) \suchthat U(u) \neq U(\south{u})\}\) are equal.
	\begin{itemize}[left=0pt]
		\item On the one hand, if \(\ind{c, d, a} = \ind{c, d, b}\),
			then \(\pp[r]{i}(T) = 0\) by \cref{item:pp-cda-cdb}.
			Then, it also holds that \(\ind{d, c, b} = \ind{d, c, a}\),
			and hence \(\pp[r]{i}(U) = 0\) by \cref{item:pp-cda-cdb}.
		\item On the other hand, if \(\ind{c, d, a}=1-\ind{c, d, b} = \ind{d, c, b}\),
			then \(\pp[r]{i}(T) = 1\) by \cref{item:pp-cda-dcb}.
			Then, it also holds that \(\ind{d, c, b} = \ind{c, d, a}\),
			and hence \(\pp[r]{i}(U) = 1\) by \cref{item:pp-cda-dcb}.
			For the (co)inversion number, we have
			\(\coinv[r, r + 1](T) - \coinv[r, r + 1](U) = \ind{d, c, b} - \ind{c, d, a} = 0\).
	\end{itemize}
	Thus, in both cases for \(\ind{c,d,a}\),
	\Cref{eq:wtqt-prob-balanced-r} holds for \(|\{a,b,c,d\}| = 4\).

	Next, suppose \(|\{a, b, c, d\}| = 3\).
	Since \(b\neq c\), we either have \(b=d\) or \(a=c\).
	Without loss of generality, assume \(b=d\),
	since swapping \(T\) and \(U\) gives the other case.
	By \cref{item:pp-bd}, we have
	\(\pp[r]{i}(T) = 0\).
	Moreover,
	\begin{equation*}
		\pp[r]{i}(U)
		=
		t^{1 - \ind{c, b, a}}
		\frac{(1 - q^{\ell+1}t^{A+1})}{1 - q^{\ell+1}t^{A+2}},
	\end{equation*}
	where \(A = \arm(i+1, r+1)\), by \cref{item:pp-ac}.
	\(\coinv[r, r + 1](T) - \coinv[r, r + 1](U) = \ind{b, c, b} - \ind{c, b, a} = 1 - \ind{c, b, a}\).
	Finally, the ratio
	\begin{equation*}
		\left(
			\prod_{\substack{u \in \dg[r + 1]{\alpha} \\ T(u) \neq T(\south{u})}}
			\frac{1 - t}{1 - q^{\leg(u)+1} t^{ \arm(u)+1}}
		\right)
		\big/
		\left(
			\prod_{\substack{u \in \dg[r + 1]{\alpha} \\ U(u) \neq U(\south{u})}}
			\frac{1 - t}{1 - q^{\leg(u)+1} t^{\arm(u)+1}}
		\right)
	\end{equation*}
	only differs in the boxes \((i,r+1)\) and \((i+1,r+1)\), and thus equals
	\begin{equation*}
		\left(
		\frac{1-t}{1 - q^{\ell+1}t^{A+2}}
		\right)
		\big/
		\left(
		\frac{1-t}{1 - q^{\ell+1}t^{A+1}}
		\right)
		= \frac{1 - q^{\ell+1}t^{A+1}}{1 - q^{\ell+1}t^{A+2}},
	\end{equation*}
	where \(A = \arm(i+1, r+1)\).
	Therefore, we compute
	\begin{equation*}
		\frac{\wtqt[r](T)}{\wtqt[r](U)} =
		t^{1 - \ind{c, b, a}} \frac{1 - q^{\ell+1}t^{A+1}}{1 - q^{\ell+1}t^{A+2}}
		= \pp[r]{i}(T, U),
	\end{equation*}
	confirming \Cref{eq:wtqt-prob-balanced-r}.

	Finally, if \(|\{a, b, c, d\}| = 2\),
	then \(a = c\) and \(b = d\).
	Hence, \(\pp[r]{i}(T) = \pp[r]{i}(U) = 1\), by \cref{item:pp-ac-bd}.
	For the (co)inversion number, we have
	\(\coinv[r, r + 1](T) - \coinv[r, r + 1](U) = \ind{b, a, b} - \ind{a, b, a} = 1 - 1 = 0\).
	Moreover, the sets \(\{u \in \Arm(i+1, r+1) \suchthat T(u) \neq T(\south{u})\}\)
	and \(\{u \in \Arm(i+1, r+1) \suchthat U(u) \neq U(\south{u})\}\) are equal.
	Thus, \Cref{eq:wtqt-prob-balanced-r} holds in this final case, as well.
\end{proof}

\begin{lemma} \label{lemma:wtqt-prob-balanced-h}
	Let \(T \in \NAF(\alpha, \sigma)\) be a non-attacking filling.
	Let \(U = \sswap[0, h]{i}(T)\) for some \(h \in \interval{0, \alpha_i}\).
	Then,
	\begin{equation} \label{eq:wtqt-prob-balanced-h}
		\wtqt[h](U)\left(1 - \pp[h]{i}(U)\right) = \wtqt[h](T)\left(1 - \pp[h]{i}(T)\right).
	\end{equation}
\end{lemma}

\begin{proof}
	If \(h = \alpha_i\),
	then \(\wtqt[\alpha_i](T) = \wtqt[\alpha_i](U) = 1\)
	and \(1 - \pp[\alpha_i]{i}(T) = 1 - \pp[\alpha_i]{i}(U) = 1\),
	so \Cref{eq:wtqt-prob-balanced-h} holds trivially.
	Assume that \(h < \alpha_i\).
	Then
	\begin{align*}
		c & = T(i, h+1)=U(i,h+1), &
		d & = T(i+1, h+1)=U(i+1,h+1), \\
		a & = T(i, h)=U(i+1,h), \text{ and } &
		b & = T(i+1, h)=U(i,h).
	\end{align*}
	Since \(U\) is non-attacking, we have \(b \neq d\).
	We note that
	\begin{equation*}
		\maj[h + 1](T) - \maj[h + 1](U) = \ind{c, a} + \ind{d, b} - \ind{c, b} - \ind{d, a}.
	\end{equation*}
	Moreover, we have
	\begin{align*}
		\coinv[h, h + 1](T) - \coinv[h, h + 1](U) & = \inv[h, h + 1](U) - \inv[h, h + 1](T)\\
		& = \ind{c, d, b} - \ind{c, d, a} +
		\sum_{\substack{u \in \Arm(i+1, h+1) \\ k = U(u)}}
		\left( \ind{c, k, b} + \ind{d,k, a} - \ind{c,k, a} - \ind{d,k, b} \right),
	\end{align*}
	which simplifies to
	\( (A+1) \left( \ind{c, a} + \ind{d, b} - \ind{c, b} - \ind{d, a} \right)\)
	using \cref{lemma:useful-ind-five-variables}, where \(A = \arm(i+1, h+1)\).

	We show that \Cref{eq:wtqt-prob-balanced-h} holds
	for all possible cases for \((a,b,c,d)\) with \(b\neq d\).

	First, suppose \(|\{a, b, c, d\}| = 4\).
	Then, the sets \(\{u \in \Arm(i+1, h+1) \suchthat T(u) \neq T(\south{u})\}\)
	and \(\{u \in \Arm(i+1, h+1) \suchthat U(u) \neq U(\south{u})\}\) are equal.
	\begin{itemize}[left=0pt]
		\item On the one hand, if \(\ind{c, d, a} = \ind{c, d, b}\),
			then \(1 - \pp[h]{i}(T) = 1\) by \cref{item:pp-cda-cdb}.
			Then, it also holds that \(\ind{d, c, b} = \ind{d, c, a}\),
			and hence \(1 - \pp[h]{i}(U) = 1\) by \cref{item:pp-cda-cdb}.
			It also holds that
			\begin{equation*}
				\ind{c, a} + \ind{d, b} - \ind{c, b} - \ind{d, a} = 0,
			\end{equation*}
			therefore, \(\maj[h + 1](T) = \maj[h + 1](U)\)
			and \(\coinv[h, h + 1](T) = \coinv[h, h + 1](U)\).
		\item On the other hand, if \(\ind{c, d, a}=1-\ind{c, d, b} = \ind{d, c, b}\),
			then \(1 - \pp[h]{i}(T) = 0\) by \cref{item:pp-cda-dcb}.
			Then, it also holds that \(\ind{d, c, b} = \ind{c, d, a}\),
			and hence \(1 - \pp[h]{i}(U) = 0\) by \cref{item:pp-cda-dcb}.
	\end{itemize}
	Thus, in both cases for \(\ind{c,d,a}\), \Cref{eq:wtqt-prob-balanced-h} holds for \(|\{a,b,c,d\}|=4\).

	Now suppose that \(|\{a, b, c, d\}| = 3\).
	Since \(b\neq d\), without loss of generality we may assume \(b=c\).
	By \cref{item:pp-bc}, we have
	\(1 - \pp[h]{i}(T) = 1\).
	Moreover,
	\begin{equation*}
		1 - \pp[h]{i}(U) = (q^{L-h}t^{A+1})^{1 - \ind{d, b, a}} \frac{1 - t}{1 - q^{L-h}t^{A+2}},
	\end{equation*}
	where \(A = \arm(i+1, h+1)\), by \cref{item:pp-ac}.
	It also holds that
	\begin{equation*}
		\ind{b, a} + \ind{d, b} - \ind{b, b} - \ind{d, a} = \ind{d, b, a}.
	\end{equation*}
	For the major index, we have \(\maj[r + 1](T) - \maj[r + 1](U)\) equals \( (\ell+1) \ind{d, b, a}\).
	For the coinversion number,
	we have \(\coinv[r, r + 1](T) - \coinv[r, r + 1](U)\)
	equals \((A + 1) \ind{d, b, a}\).
	Finally, we have
	\begin{equation*}
		\left(
			\prod_{\substack{u \in \dg[r + 1]{\alpha} \\ T(u) \neq T(\south{u})}}
			\frac{1 - t}{1 - q^{\leg(u)+1} t^{ \arm(u)+1}}
		\right)
		\big/
		\left(
			\prod_{\substack{u \in \dg[r + 1]{\alpha} \\ U(u) \neq U(\south{u})}}
			\frac{1 - t}{1 - q^{\leg(u)+1} t^{\arm(u)+1}}
		\right)
		= \frac{(1 - t)}{1 - q^{\ell+1}t^{A+1}},
	\end{equation*}
	since \(u=(i, r+1)\) is the only relevant box.
	Therefore, we compute
	\begin{equation*}
		\frac{\wtqt[r](T)}{\wtqt[r](U)} =
		q^{(\ell+1)\ind{d, b, a}} t^{(A+1)\ind{d, b, a}} \frac{(1 - t)}{1 - q^{\ell+1}t^{A+1}}
		= \pp[r]{i}(T, U),
	\end{equation*}
	and thus \Cref{eq:wtqt-prob-balanced-h} holds for this case as well.

	Finally, the case \(|\{a, b, c, d\}| = 2\) does not happen as \(b \neq d\)
	by the non-attacking property of \(U\).
	Therefore, \Cref{eq:wtqt-prob-balanced-h} holds in all cases.
\end{proof}

We are now ready to prove the main result of this section, \cref{theorem:wt-prob}.

\begin{proof}[Proof of \cref{theorem:wt-prob}]
	Assume that \(U = \sswap[0, h]{i}(T)\) for some \(h \in \interval{0, \alpha_i}\).
	If such \(h\) doesn't exist, then \(\prob{i}(T, U) = \prob{i}(U, T) = 0\), and the lemma holds trivially.
	For all \(r \in \interval{h + 1, \alpha_i-1}\),
	since the \(r\)\textsuperscript{th} and \((r+1)\)\textsuperscript{st} rows of \(T\) and \(U\) are equal,
	we have
	\begin{equation} \label{eq:wtqt-prob-balanced-r-large}
		\wtqt[r](T) = \wtqt[r](U)
	\end{equation}
	By \cref{lemma:wtqt-prob-balanced-r}, for all \(r \in \interval{0, h-1}\), we have
	\begin{equation} \label{eq:wtqt-prob-balanced-r-repeated}
		\wtqt[r](U)\pp[r]{i}(U) = \wtqt[r](T)\pp[r]{i}(T),
	\end{equation}
	and by \cref{lemma:wtqt-prob-balanced-h}, we have
	\begin{equation} \label{eq:wtqt-prob-balanced-h-repeated}
		\wtqt[h](U)\left(1 - \pp[h]{i}(U)\right) = \wtqt[h](T)\left(1 - \pp[h]{i}(T)\right).
	\end{equation}
	Multiplying
	\Cref{eq:wtqt-prob-balanced-r-large} for each \(r \in \interval{h + 1, \alpha_i-1}\),
	\Cref{eq:wtqt-prob-balanced-r-repeated} for all \(r \in \interval{0, h-1}\), and
	\Cref{eq:wtqt-prob-balanced-h-repeated},
	we obtain
	\begin{equation*}
		\left( \prod_{r = 0}^{\alpha_i-1} \wtqt[r](U) \right)
		\left( \prod_{r = 0}^{h-1} \pp[r]{i}(U) \right)
		\left( 1 - \pp[h]{i}(U) \right)
		=
		\left( \prod_{r = 0}^{\alpha_i-1} \wtqt[r](T) \right)
		\left( \prod_{r = 0}^{h-1} \pp[r]{i}(T) \right)
		\left( 1 - \pp[h]{i}(T) \right).
	\end{equation*}
	By \Cref{eq:rthcomponent}, and \cref{definition:prob}, we obtain
	\begin{equation*}
		\wtqt(U)\prob{i}(U, T) = \wtqt(T)\prob{i}(T, U). \qedhere
	\end{equation*}
\end{proof}

Recall that \(\NAF(\alpha, \sigma, \beta)\) denotes the set of non-attacking fillings of shape \(\alpha\) with content \(\beta\).
Since \(\prob{i}(T, U) = 0\) whenever \(\mathbf{x}^T \neq \mathbf{x}^U\),
we may restrict the map in \cref{theorem:wt-prob}
to obtain a probabilistic bijection between
\(\NAF(\alpha, \sigma, \beta)\) and \(\NAF(\alpha, \sigma s_i, \beta)\).

\begin{corollary} \label{cor:wt-prob}
	The pair of maps
	% \begin{gather*}
	% 	{\prob{i}} \colon \NAF(\alpha, \sigma, \beta) \times \NAF(\alpha, \sigma s_i, \beta) \to \rationals(q, t) \\
	% 	{\prob{i}} \colon \NAF(\alpha, \sigma s_i, \beta) \times \NAF(\alpha, \sigma, \beta) \to \rationals(q, t)
	% \end{gather*}
	defines a probabilistic bijection between
	\(\NAF(\alpha, \sigma, \beta)\) and \(\NAF(\alpha, \sigma s_i, \beta)\)
	with respect to the weight function \(\wtqt\).
\end{corollary}

\begin{example}
	Let \(n = 4\),
	\(\alpha = \composition{2, 2, 0, 1}\),
	\(\sigma = \window{3, 1, 2, 4}\),
	\(\beta = \composition{1, 0, 2, 2}\).
	Let \(i = 1\) and note that \(\alpha_1 = \alpha_2 = 2\).
	Moreover, note that \(\sigma s_i = \window{1, 3, 2, 4}\).
	\cref{cor:wt-prob} implies that the maps
	\( \prob{1} \colon \NAF(\alpha, \sigma, \beta) \times \NAF(\alpha, \sigma s_1, \beta) \to \rationals(q, t)\)
	and
	\( \prob{1} \colon \NAF(\alpha, \sigma s_1, \beta) \times \NAF(\alpha, \sigma, \beta) \to \rationals(q, t)\)
	form a probabilistic bijection.

	The set
	\(\NAF(\alpha, \sigma, \beta) = \NAF(\composition{2, 2, 0, 1}, \window{3, 1, 2, 4}, \composition{1, 0, 2, 2})\)
	consists of three fillings,
	\(\textcolor{red!50!black}{T_1}\),
	\(\textcolor{red!50!black}{T_2}\), and
	\(\textcolor{red!50!black}{T_3}\),
	and the set
	\(\NAF(\alpha, \sigma s_i, \beta) = \NAF(\composition{2, 2, 0, 1}, \window{1, 3, 2, 4}, \composition{1, 0, 2, 2})\) 
	consists of two fillings,
	\(\textcolor{blue!50!black}{U_1}\) and
	\(\textcolor{blue!50!black}{U_2}\),
	shown in the left and right sides of \cref{fig:prob-2201-3124-1022}, respectively.
	\cref{fig:prob-2201-3124-1022} shows the values of
	\(\prob{1}(T, U)\) and \(\prob{1}(U, T)\) for each pair of fillings.

	\begin{figure}[htbp]
		\centering
		\begin{tikzpicture}
			\node[red!50!black]  (T1) at (-4,  3) {
				\(T_1 =
				\begin{ytableau}
					2 & 4 \\
					1 & 2 & \none & 4 \\
					3 & 1 & 2 & 4
				\end{ytableau}\)
			};
			\node[red!50!black]  (T2) at (-4,  0) {
				\(T_2 =
				\begin{ytableau}
					2 & 4 \\
					2 & 1 & \none & 4 \\
					3 & 1 & 2 & 4
				\end{ytableau}\)
			};
			\node[red!50!black]  (T3) at (-4, -3) {
				\(T_3 =
				\begin{ytableau}
					4 & 2 \\
					1 & 2 & \none & 4 \\
					3 & 1 & 2 & 4
				\end{ytableau}\)
			};
			\node[blue!50!black] (U1) at ( 4,  1.5) {
				\(\begin{ytableau}
					2 & 4 \\
					1 & 2 & \none & 4 \\
					1 & 3 & 2 & 4
				\end{ytableau} = U_1\)
			};
			\node[blue!50!black] (U2) at ( 4, -1.5) {
				\(\begin{ytableau}
					4 & 2 \\
					1 & 2 & \none & 4 \\
					1 & 3 & 2 & 4
				\end{ytableau} = U_2\)
			};
			% Weights
			\node[red!50!black, below] at (T1.south) {
				\(\scriptstyle \wtqt(T_1) =
				\frac{q^{4} {(1 - t)}^{4} t^{3}}{{(1 - q^{2} t^{3})} {(1 - q^{2} t^{2})} {(1 - q t^{2})} {(1 - q t)}}\)
			};
			\node[red!50!black, below] at (T2.south) {
				\(\scriptstyle \wtqt(T_2) =
				\frac{q {(1 - t)}^{2}}{{(1 - q^{2} t^{3})} {(1 - q t)}}\)
			};
			\node[red!50!black, below] at (T3.south) {
				\(\scriptstyle \wtqt(T_3) =
				\frac{q^{3} t^{2} {(1 - t)}^{3} }{{(1 - q^{2} t^{3})} {(1 - q^{2} t^{2})} {(1 - q t^{2})}}\)
			};
			\node[blue!50!black, below] at (U1.south) {
				\(\scriptstyle \wtqt(U_1) =
				\frac{q^{2} {(1 - t)}^{3} t}{{(1 - q^{2} t^{2})} {(1 - q t^{2})} {(1 - q t)}}\)
			};
			\node[blue!50!black, below] at (U2.south) {
				\(\scriptstyle \wtqt(U_2) =
				\frac{q {(1 - t)}^{2}}{{(1 - q^{2} t^{2})} {(1 - q t^{2})}}\)
			};
			% Arrows
			\draw[-{Latex}, red!50!black] (T1) 
				edge[bend left=15] node[sloped, fill=white]{\(1\)} (U1);
			\draw[-{Latex}, blue!50!black] (U1) 
				edge[bend left=0] node[sloped, fill=white]{\(\tfrac{q^2t^2(1-t)}{1-q^2t^3}\)} (T1);
			\draw[-{Latex}, red!50!black] (T2) 
				edge[bend left=10] node[sloped, fill=white]{\(\frac{qt(1-t)}{1-qt^2}\)} (U1);
			\draw[-{Latex}, blue!50!black] (U1) 
				edge[bend left=6] node[sloped, fill=white]{\(\frac{1-q^2t^2}{1-q^2t^3}\)} (T2);
			\draw[-{Latex}, red!50!black] (T2) 
				edge[bend left=6] node[sloped, fill=white]{\(\frac{qt-1}{1-qt^2}\)} (U2);
			\draw[-{Latex}, blue!50!black] (U2) 
				edge[bend left=10] node[sloped, fill=white]{\(\frac{1-q^2t^2}{1-q^2t^3}\)} (T2);
			\draw[-{Latex}, red!50!black] (T3) 
				edge[bend left=0] node[sloped, fill=white]{\(1\)} (U2);
			\draw[-{Latex}, blue!50!black] (U2) 
				edge[bend left=15] node[sloped, fill=white]{\(\tfrac{q^2t^2(1-t)}{1-q^2t^3}\)} (T3);
		\end{tikzpicture}
		\caption{
			For each \(T \in \NAF(\composition{2, 2, 0, 1}, \window{3, 1, 2, 4}, \composition{1, 0, 2, 2})\)
			(right, in red),
			and each \(U \in \NAF(\composition{2, 2, 0, 1}, \window{1, 3, 2, 4}, \composition{1, 0, 2, 2})\)
			(left, in blue),
			the label of the red arrow from \(T\) to \(U\) is the value of \(\prob{1}(T, U)\),
			and the label of the blue arrow from \(U\) to \(T\) is the value of \(\prob{1}(U, T)\).
			The lack of an arrow indicates that the value is zero.
			Below each filling, we show the value of \(\wtqt\).
		}
		\label{fig:prob-2201-3124-1022}
	\end{figure}

	The fact that the maps form a probabilistic bijection can be checked
	by computing that, for each
	\(T \in \{\textcolor{red!50!black}{T_1}, \textcolor{red!50!black}{T_2}, \textcolor{red!50!black}{T_3}\}\)
	and each
	\(U \in \{\textcolor{blue!50!black}{U_1}, \textcolor{blue!50!black}{U_2}\}\),
	we have
	\begin{equation*}
		\prob{1}(T, U) \wtqt(T) = \prob{1}(U, T) \wtqt(U).
	\end{equation*}
	Moreover, using the fact that there exists a probabilistic bijection between
	\(\{\textcolor{red!50!black}{T_1}, \textcolor{red!50!black}{T_2}, \textcolor{red!50!black}{T_3}\}\)
	and
	\(\{\textcolor{blue!50!black}{U_1}, \textcolor{blue!50!black}{U_2}\}\),
	\cref{theorem:wt-prob} implies that their \(q, t\)-weight generating functions are equal,
	that is,
	\begin{equation*}
		\wtqt(\textcolor{red!50!black}{T_1})
		+ \wtqt(\textcolor{red!50!black}{T_2})
		+ \wtqt(\textcolor{red!50!black}{T_3})
		= \wtqt(\textcolor{blue!50!black}{U_1})
		+ \wtqt(\textcolor{blue!50!black}{U_2}).
	\end{equation*}
	This equality can be checked from the known values of \(\wtqt\) in \cref{fig:prob-2201-3124-1022}.
	Note that the left-hand side is the coefficient
	\([\mathbf{x}^{\composition{1, 0, 2, 2}}] E_{\composition{2, 2, 0, 1}}^{\window{3, 1, 2, 4}}(\mathbf{x}; q, t)\)
	and the right-hand side is the coefficient
	\([\mathbf{x}^{\composition{1, 0, 2, 2}}] E_{\composition{2, 2, 0, 1}}^{\window{1, 3, 2, 4}}(\mathbf{x}; q, t)\).
\end{example}

\subsection{Bijective proof of \texorpdfstring{\cref{theorem:main}}{symmetry theorem}}
\label{sec:proof}

We now have the tools to give a bijective proof for the main result of this chapter,
\cref{theorem:main}, a symmetry theorem for the permuted-basement Macdonald polynomials.
Let's recall the statement.

\main*

\begin{proof}
	We write
	\begin{equation*}
		E_{\alpha}^{\sigma}(\mathbf{x}; q, t)
		= \sum_{T \in \NAF(\alpha, \sigma)} \mkern-9mu \wt(T)
		= \sum_{\text{compositions }\beta} \mkern-9mu x^\beta
			\left(\sum_{T \in \NAF(\alpha, \sigma, \beta)} \mkern-9mu \wtqt(T)\right).
	\end{equation*}
	For each composition \(\beta = \composition{\beta_1, \ldots, \beta_n}\),
	\cref{cor:wt-prob} implies that there exists a probabilistic bijection between
	\(\NAF(\alpha, \sigma, \beta)\) and \(\NAF(\alpha, \sigma s_i, \beta)\)
	with respect to \(\wtqt\).
	Hence, \cref{proposition:equal-weight-sum} implies that
	\begin{equation*}
		[x^\beta] E_{\alpha}^{\sigma}(\mathbf{x}; q, t)
		= \sum_{T \in \NAF(\alpha, \sigma, \beta)} \mkern-9mu \wtqt(T)
		= \sum_{U \in \NAF(\alpha, \sigma s_i, \beta)} \mkern-9mu \wtqt(U)
		= [x^\beta] E_{\alpha}^{\sigma s_i}(\mathbf{x}; q, t),
	\end{equation*}
	and the result follows.
\end{proof}

Given a composition \(\alpha=\composition{\alpha_1,\ldots,\alpha_n}\),
let \(\mathscr{C}_\alpha\) be the subgroup of \(\sym{n}(\alpha)\)
generated by the transpositions \(\{s_i \suchthat \alpha_i = \alpha_{i+1}\}\).
Writing \(\alpha = p_1^{m_1} p_2^{m_2} \cdots p_k^{m_k}\),
we have that
\begin{equation*}
	\mathscr{C}_\alpha =
	\sym{\interval{1, m_1}} \times
	\sym{\interval{m_1+1, m_1+m_2}} \times
	\cdots \times
	\sym{\interval{m_1+m_2+\cdots+m_{k-1}+1, n}}
	\subseteq \sym{n}(\alpha).
\end{equation*}

For example, if \(\alpha =\composition{1, 1, 2, 2, 1, 3} = 1^2 2^2 1^1 3^1\),
then \(\mathscr{C}_\alpha\) is the subgroup of \(\sym{6}(\alpha)\) generated by
the transpositions \(s_1\) and \(s_3\):
\begin{equation*}
	\mathscr{C}_\alpha =
	\{
		\window{1, 2, 3, 4, 5, 6},
		\window{2, 1, 3, 4, 5, 6},
		\window{1, 2, 4, 3, 5, 6},
		\window{4, 1, 4, 3, 5, 6}
	\}.
\end{equation*}
The group \(\sym{6}(\alpha)\) is generated by \(\{s_1,s_3,s_{2,5}\}\),
where \(s_{2,5}=s_2s_3s_4s_3s_2\).
For instance, \(\window{5, 2, 3, 4, 1, 6}\in\sym{n}(\alpha)\setminus \mathscr{C}_\alpha\).

\begin{corollary}
	\label{corollary:C-alpha-symmetry}
	Let \(\alpha=\composition{\alpha_1,\ldots,\alpha_n}\) be a composition.
	Then, if \(\sigma, \tau \in \sym{n}\) are such that
	\(\sigma^{-1} \tau \in \mathscr{C}_\alpha\), then
	\begin{equation*}
		E_{\alpha}^{\sigma}(\mathbf{x}; q, t) = E_{\alpha}^{\tau}(\mathbf{x}; q, t).
	\end{equation*}
\end{corollary}

\begin{proof}
	By induction,
	it suffices to prove the statement for \(\sigma^{-1} \tau = s_i\)
	for some \(i\) such that \(\alpha_i = \alpha_{i+1}\).
	Then, the result follows from \cref{theorem:main}.
\end{proof}

If \(\alpha = p_1^{m_1} p_2^{m_2} \cdots p_k^{m_k}\),
then \(\mathscr{C}_\alpha = \sym{n}(\alpha)\)
if and only if \(p_1, p_2, \ldots, p_k\) are pairwise distinct.
In particular,
this is the case when \(\alpha\) is a partition or an antipartition.

\begin{corollary} \label{corollary:partition-antipartition-symmetry}
	Let \(\alpha = \composition{\alpha_1, \ldots, \alpha_n}\)
	be a partition or an antipartition.
	If \(\sigma, \tau \in \sym{n}\) are such that
	\(\sigma \permact \alpha = \tau \permact \alpha\),
	then
	\begin{equation*}
		E_{\alpha}^{\sigma}(\mathbf{x}; q, t) = E_{\alpha}^{\tau}(\mathbf{x}; q, t).
	\end{equation*}
\end{corollary}

\begin{proof}
	Since \(\sigma^{-1} \tau \in \sym{n}(\alpha)\),
	when \(\alpha\) is a partition or an antipartition,
	\(\sigma^{-1} \tau \in \mathscr{C}_\alpha\),
	and \cref{corollary:C-alpha-symmetry} applies.
\end{proof}

\section{Applications}\label{sec:applications}

Immediate applications of \cref{theorem:main} include the removal of certain assumptions from theorems in \cite{Ale19} and \cite{CMW22}.

\subsection{\texorpdfstring{\(t\)}{t}-atom polynomials and \texorpdfstring{\(t\)}{t}-key polynomials}
\label{sec:t-atoms}

Fix a composition \(\alpha = \composition{\alpha_1, \ldots, \alpha_n}\)
and a permutation \(\sigma \in \sym{n}\).
The \vocab{permuted basement \(t\)-atom polynomial of \(\alpha\) with basement \(\sigma\)},
first introduced in \cite{HLMvW11} and further studied in \cite{AS19}, is defined as
\begin{equation*}
	\atompoly_{\alpha}^{\sigma}(\mathbf{x}; t) = E_{\alpha}^{\sigma}(\mathbf{x}; 0, t).
\end{equation*}
Similarly, the \vocab{\(t\)-atom polynomial} \(\atompoly_{\alpha}(\mathbf{x}; t)\)
and the \vocab{\(t\)-key polynomial} \(\keypoly_{\alpha}(\mathbf{x}; t)\) are defined as
\begin{equation*}
	\atompoly_{\alpha}(\mathbf{x}; t) = \atompoly_{\alpha}^{\id}(\mathbf{x}; t)
	\qquad \text{and} \qquad
	\keypoly_{\alpha}(\mathbf{x}; t) = \atompoly_{\alpha}^{w_0}(\mathbf{x}; t).
\end{equation*}

\begin{proposition}[inferred from {\cite[Proposition~27]{Ale19}}]
	\label{proposition:atom-alpha-inc-alpha}
	Let \(\alpha = \composition{\alpha_1, \ldots, \alpha_n}\) be a composition.
	Let \(\sigma \in \sym{n}\) be the shortest permutation such that
	\(\sigma \permact \inc(\alpha) = \alpha\), and
	let \(\pi \in \sym{n}\) be the shortest permutation such that
	\(\pi \permact \dec(\alpha) = \alpha\).
	Then,\todo[inline]{Review the details of this proposition.}
	\begin{equation*}
		\atompoly_{\alpha}(\mathbf{x}; t) = \atompoly_{\inc(\alpha)}^{\sigma}(\mathbf{x}; t)\qquad\text{and}\qquad
		\keypoly_{\alpha}(\mathbf{x}; t) = \atompoly_{\dec(\alpha)}^{\pi}(\mathbf{x}; t).
	\end{equation*}
\end{proposition}

Using \cref{corollary:partition-antipartition-symmetry},
we can remove the length restriction on \(\sigma\) and \(\pi\) above.

\begin{corollary} \label{lemma:atom-alpha-inc-alpha}
	Let \(\alpha = \composition{\alpha_1, \ldots, \alpha_n}\) be a composition.
	Let \(\tau \in \sym{n}\) be any permutation such that
	\(\tau \permact \inc(\alpha) = \alpha\), and
	let \(\pi \in \sym{n}\) be any permutation such that
	\(\pi \permact \dec(\alpha) = \alpha\).
	Then,
	\begin{equation*}
		\atompoly_{\alpha}(\mathbf{x}; t) = \atompoly_{\inc(\alpha)}^{\tau}(\mathbf{x}; t)
		\qquad\text{and}\qquad
		\keypoly_{\alpha}(\mathbf{x}; t) = \atompoly_{\dec(\alpha)}^{\pi}(\mathbf{x}; t).
	\end{equation*}
\end{corollary}

\begin{proof}
	Let \(\sigma \in \sym{n}\) be the shortest permutation such that \(\sigma \permact \inc(\alpha) = \alpha\).
	By Proposition \ref{proposition:atom-alpha-inc-alpha}, we have
	\(
		\atompoly_{\alpha}(\mathbf{x}; t)
		= \atompoly_{\inc(\alpha)}^{\sigma}(\mathbf{x}; t)
		= E_{\inc(\alpha)}^{\sigma}(\mathbf{x}; 0, t)
	\).
	Since \(\tau \permact \inc(\alpha) = \sigma \permact \inc(\alpha)\),
	\cref{corollary:partition-antipartition-symmetry} implies
	\(
		E_{\inc(\alpha)}^{\sigma}(\mathbf{x}; 0, t)
		= E_{\inc(\alpha)}^{\tau}(\mathbf{x}; 0, t)
		= \atompoly_{\inc(\alpha)}^{\tau}(\mathbf{x}; t)
	\).

	The proof for \(\keypoly_{\alpha}\) is analogous, and is omitted.
\end{proof}

\subsection{ASEP polynomials}\label{sec:mlqs}

In \cite{CMW22}, the authors introduce the ASEP polynomial \(F_\alpha(\mathbf{x}; q, t)\),
which are weight-generating functions for multiline queues with bottom row \(\alpha\).
A key result from their work is the following proposition.

\begin{proposition}[{\cite[Proposition~3.1]{CMW22}}] \label{proposition:multiline-queues}
	Let \(\alpha = \composition{\alpha_1, \ldots, \alpha_n}\) be a composition.
	Let \(\sigma \in \sym{n}\) be the longest permutation such that
	\(\sigma^{-1} \permact \alpha = \inc(\alpha)\).
	Then,
	\begin{equation*}
		F_{\alpha}(\mathbf{x}; q, t) = E_{\inc(\alpha)}^{\sigma}(\mathbf{x}; q, t).
	\end{equation*}
\end{proposition}

By \cref{corollary:partition-antipartition-symmetry},
we can remove the length requirement on \(\sigma\).

\begin{corollary}
	\label{lemma:multiline-queues}
	Let \(\alpha=\composition{\alpha_1,\ldots,\alpha_n}\) be a composition, and let
	\(\tau \in \sym{n}\) such that \(\tau^{-1} \permact \alpha = \inc(\alpha)\).
	Then,
	\begin{equation*}
		F_{\alpha}(\mathbf{x}; q, t) = E_{\inc(\alpha)}^{\tau}(\mathbf{x}; q, t).
	\end{equation*}
\end{corollary}

This allows us to realize \(t\)-atom polynomials as specializations of ASEP polynomials.

\begin{corollary}
	Let \(\alpha = \composition{\alpha_1, \ldots, \alpha_n}\) be a composition.
	Then,
	\begin{equation*}
		F_{\alpha}(\mathbf{x}; 0, t) = \atompoly_{\alpha}(\mathbf{x}; t).
	\end{equation*}
\end{corollary}

\begin{proof}
	Let \(\tau \in \sym{n}\) such that \(\tau^{-1} \permact \alpha = \inc(\alpha)\).
	From \cref{lemma:multiline-queues}
	and \cref{lemma:atom-alpha-inc-alpha}, we have
	\(
		F_{\alpha}(\mathbf{x}; 0, t)
		= E_{\inc(\alpha)}^{\tau}(\mathbf{x}; 0, t)
		= \atompoly_{\inc(\alpha)}^{\tau}(\mathbf{x}; t)
		= \atompoly_{\alpha}(\mathbf{x}; t)
	\).
\end{proof}

Finally, for a partition \(\lambda=\composition{\lambda_1,\ldots,\lambda_n}\), it was observed in \cite{CMW22} that  
\begin{equation}
	\label{eq:P in Es}
	P_{\lambda}(\mathbf{x};q,t)
	=
	\sum_{\mu\in S_n\permact \lambda}
	E_{\inc(\lambda)}^{\sigma_\mu}(\mathbf{x};q,t),
\end{equation}
where \(S_n\permact \lambda\) is the set of rearrangements of \(\lambda\), \(\inc(\lambda)=(\lambda_n,\ldots,\lambda_1)\), and \(\sigma_\mu\) is the longest permutation that rearranges \(\inc(\lambda)\) into \(\mu\).

\Cref{corollary:partition-antipartition-symmetry} allows us to rewrite \Cref{eq:P in Es} without the length restriction on \(\sigma_\mu\).
\begin{corollary}
	Let \(\lambda=\composition{\lambda_1,\ldots,\lambda_n}\) be a partition. Then the symmetric Macdonald polynomial satisfies
	\begin{equation*}
		P_\lambda(\mathbf{x};q,t)
		=\sum_{\mu\in S_n\cdot\lambda}
		E_{\inc(\lambda)}^{\sigma_\mu}(\mathbf{x};q,t),
	\end{equation*}
	where \(\sigma_\mu\in S_n\) is any permutation such that \(\sigma_\mu\permact \inc(\lambda)=\mu\).
\end{corollary}